---
title: "01_Read_fix_NIFES_data"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
---

Based on code from '04_NIVA_data_with_NIFES_3.R' ("H:/Documents/seksjon 212/Indikator 2018/Analyse")  

## 1. Packages  
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)

# save_data <- TRUE
save_data <- FALSE
```

## 2. Data  
```{r}
dat1 <- read_excel("Input_data/Data kysttorsk NH nov. 2018.xlsx", sheet = "Metaller i torskefilet")
dat2 <- read_excel("Input_data/Data kysttorsk NH nov. 2018.xlsx", sheet = "Metaller og POPs i torskelever", skip = 2)

colnames(dat1)
# cat("\n")
# colnames(dat2)

# Check uniqueness of Jnr
nrow(dat1) == length(unique(dat1$Jnr))
nrow(dat2) == length(unique(dat2$Jnr))
```

## 3. Put dat1 on tidy ('tall') format
### Get column numbers
```{r}
# colnames(dat1) %>% dput()

colnumber_unit <- which(substr(colnames(dat1),1,5) %in% "Enhet")
colnumber_conc <- colnumber_unit - 1
colnumber_meta <- 1:(colnumber_conc[1]-1)
```

### Make concentration data
```{r}
dat1_conc <- dat1[c(colnumber_meta, colnumber_conc)] %>%
  gather("Parameter", "Conc_chr", `Tørrstoff %`:Ni) %>%                            # Note hard-coded column names
  as.data.frame()

# Check first and second character of value. The "<" is always in the first
table(substr(dat1_conc$Conc_chr, 1, 1))
table(substr(dat1_conc$Conc_chr, 2, 2))

# Remove less-than sign and make Conc variabøe
dat1_conc$Conc_chr2 <- sub("<", "", dat1_conc$Conc_chr, fixed = TRUE)
dat1_conc$Conc <- as.numeric(dat1_conc$Conc_chr2)

# Find less-thans and make Flag variable
dat1_conc$Flag <- ifelse(substr(dat1_conc$Conc_chr, 1, 1) %in% "<", "<", "")

nrow(dat1_conc)
```

### Make unit data
```{r}
dat1_unit <- dat1[c(1, colnumber_unit)] %>%
  as.data.frame()

# Set column names (which now are "Enhet__1" etc.) to the corresponding actual name
colnames(dat1_unit)[-1] <- colnames(dat1)[colnumber_conc]

dat1_unit <- dat1_unit %>%
  gather("Parameter", "Unit", `Tørrstoff %`:Ni) %>%                               # Note hard-coded column names
  as.data.frame()

nrow(dat1_unit)
```

### Put dat1_conc and dat1_unit together
```{r}
check <- sum(dat1_conc$Jnr != dat1_unit$Jnr)
cat("Number of Jnr not equal:", sum(check), "\n")

if (!is.na(check) & check == 0){
  dat1_tall <- cbind(dat1_conc, Unit = dat1_unit$Unit, stringsAsFactors = FALSE)
}
```


## 4. Put dat2 on tidy ('tall') format
### Get column numbers
```{r}
# colnames(dat2) %>% dput()

colnumber_unit <- which(substr(colnames(dat2),1,5) %in% "Enhet")
colnumber_conc <- colnumber_unit - 1
colnumber_meta <- 1:(colnumber_conc[1]-1)

# colnames(dat2)[colnumber_meta]
# colnames(dat2)[colnumber_conc]
# colnames(dat2)[colnumber_unit]
```

### Make concentration data
```{r}
dat2_conc <- dat2[c(colnumber_meta, colnumber_conc)] %>%
  gather("Parameter", "Conc_chr", `Totalfett Etylacetat`:`Sum PBDE 7`) %>%                             # Note hard-coded column names
  as.data.frame()

# Check first and second character of value. The "<" is always in the first
table(substr(dat2_conc$Conc_chr, 1, 1))
table(substr(dat2_conc$Conc_chr, 2, 2))

# Remove less-than sign and make Conc variabøe
dat2_conc$Conc_chr2 <- sub("<", "", dat2_conc$Conc_chr, fixed = TRUE)
dat2_conc$Conc <- as.numeric(dat2_conc$Conc_chr2)

# Find less-thans and make Flag variable
dat2_conc$Flag <- ifelse(substr(dat2_conc$Conc_chr, 1, 1) %in% "<", "<", "")

nrow(dat2_conc)
```

### Make unit data
```{r}
dat2_unit <- dat2[c(1, colnumber_unit)] %>%
  as.data.frame()

# Set column names (which now are "Enhet__1" etc.) to the corresponding actual name
colnames(dat2_unit)[-1] <- colnames(dat2)[colnumber_conc]

dat2_unit <- dat2_unit %>%
  gather("Parameter", "Unit", `Totalfett Etylacetat`:`Sum PBDE 7`) %>%                             # Note hard-coded column names
  as.data.frame()

nrow(dat2_unit)
```

### Put dat2_conc and dat2_unit together
```{r}
check <- sum(dat2_conc$Jnr != dat2_unit$Jnr)
cat("Number of Jnr not equal:", sum(check), "\n")

if (!is.na(check) & check == 0){
  dat2_tall <- cbind(dat2_conc, Unit = dat2_unit$Unit, stringsAsFactors = FALSE)
}
```

## 5. Put dat1_tall and dat2_tall together
```{r}
df_nifes_cod <- bind_rows(dat1_tall, dat2_tall)
nrow(df_nifes_cod)  
```

## 6. Check data
```{r}
# Missing Conc values
sel <- is.na(df_nifes_cod$Conc); sum(sel)
df_nifes_cod[sel,] %>% select(Parameter, Conc_chr, Conc_chr2, Conc, Flag) %>% head()

# Missing Conc but unmissing Conc_chr values
sel <- is.na(df_nifes_cod$Conc) & !is.na(df_nifes_cod$Conc_chr); sum(sel)

# Check one example where second character = " "
tab <- table(substr(dat2_conc$Conc_chr, 2, 2))
names(tab)
sel <- substr(df_nifes_cod$Conc_chr,2,2) %in% " "; sum(sel)
df_nifes_cod[sel,] %>% select(Parameter, Conc_chr, Conc_chr2, Conc, Flag) %>% head()

# Check one example where second character = ""
sel <- substr(df_nifes_cod$Conc_chr,2,2) %in% ""; sum(sel)
df_nifes_cod[sel,] %>% select(Parameter, Conc_chr, Conc_chr2, Conc, Flag) %>% head()

# Check one example where exponent value is used (3.3911910439999998E-3)
df_nifes_cod %>% filter(Jnr %in% "2014-1753/21") %>% select(Parameter, Conc_chr, Conc)
```

## 7. Create df_nifes_cod2
### Add Lat, Long
```{r}
# df_nifes_cod %>% head(3)

df_nifes_cod <- df_nifes_cod %>%
  separate(Uttakssted, c("Lat_string","Lon_string"), sep = "[:blank:]", remove = FALSE) 

df_nifes_cod %>% select(Jnr, Uttakssted:Lon_string) %>% head(3)

# Some (ca 20%) have "N" and "E" as a part of Lat_string/Lon_string
x <- stringr::str_extract(df_nifes_cod$Lat_string, "[a-z,A-Z]")
table(addNA(x))
x <- stringr::str_extract(df_nifes_cod$Lon_string, "[a-z,A-Z]")
table(addNA(x))

# Remove letters and create LATITUDE, LONGITUDE
df_nifes_cod <- df_nifes_cod %>%
  mutate(Lat_string = sub("[a-z,A-Z]", "", Lat_string),
         Lon_string = sub("[a-z,A-Z]", "", Lon_string)) %>%
  mutate(LATITUDE = as.numeric(Lat_string),
         LONGITUDE = as.numeric(Lon_string))
```

### Select/order variables
```{r}
# Column names, replace space with underscore 
colnames(df_nifes_cod) <- gsub(" ", "_", colnames(df_nifes_cod), fixed = TRUE)

df_nifes_cod <- df_nifes_cod %>%
  select(Jnr,Merking,År,Uttaksdato,Uttakssted,LATITUDE,LONGITUDE,Uttaksområde,Art,Organ,Latinsk_navn,
         Parameter, Conc, Flag, Unit)

```

### Add sum PBDE with 6 congeners (they have only calculated "Sum PBDE 7")  
Creates df_nifes_cod2  
```{r}
df_nifes_cod_PBDE6 <- df_nifes_cod %>% 
  filter(Parameter %in% c("PBDE 28", "PBDE 47", "PBDE 99", "PBDE 100", "PBDE 126", "PBDE 153")) %>%
  group_by(Jnr,Merking,År,Uttaksdato,Uttakssted,LATITUDE,LONGITUDE,Uttaksområde,Art,Organ,Latinsk_navn,Unit) %>%
  summarise(Conc = sum(Conc, na.rm = TRUE), n = n(), BelowLOQ = sum(Flag %in% "<"), BelowLOQ_fraction = BelowLOQ/n)
summary(df_nifes_cod_PBDE6$BelowLOQ_fraction)

df_nifes_cod_PBDE6 <- df_nifes_cod_PBDE6 %>%
  mutate(Flag = ifelse(BelowLOQ_fraction >= 0.5, "<", ""))  # Set Flag to "<" if >=50% of congenerers are below LOQ
         
#
# Add "Parameter" and sort columns in same order
#
df_nifes_cod_PBDE6$Parameter <- "BDE6S"   # same name as in NIVA data
df_nifes_cod_PBDE6 <- df_nifes_cod_PBDE6[,colnames(df_nifes_cod)] %>% as.data.frame()

#
# Add to dataset, create 'df_nifes_cod2'
#
df_nifes_cod2 <- rbind(df_nifes_cod, df_nifes_cod_PBDE6, stringsAsFactors = FALSE)
nrow(df_nifes_cod)  # 36569
nrow(df_nifes_cod2) # 36946

```


### Measurement year = 1.april to 31. March
```{r}
df_nifes_cod2$Time <- lubridate::dmy(df_nifes_cod2$Uttaksdato)
df_nifes_cod2$Month <- lubridate::month(df_nifes_cod2$Time)

# Lacking Year: none; lacking date: 2450
xtabs(~addNA(År), df_nifes_cod2)
xtabs(~addNA(Month), df_nifes_cod2)
df_nifes_cod2 %>% filter(is.na(Month)) %>% select(År, Uttaksdato, Time, Month) %>% head(4)

# Set measurement year (32% are in January-March, all in År = 2010,2011,2018)
yr <- df_nifes_cod2$År
sel <- with(df_nifes_cod2, Month <= 3 & !is.na(Month))
sum(sel, na.rm = TRUE); mean(sel, na.rm = TRUE)   # 9234 = 8.2%
yr[sel] <- yr[sel]-1

# We REDEFINE Year as measurement year
df_nifes_cod2$Year <- yr

```

## 8. Some tables
```{r}
# Calender year (År) vs month
xtabs(~addNA(År) + addNA(Month), df_nifes_cod2)

# Calender year (År) vs measurement year (Year)
xtabs(~addNA(År) + addNA(Year), df_nifes_cod2)   
# All 2010,2018 data and much 2011 data are in the measurement year before

# Areas
xtabs(~addNA(Uttaksområde) + addNA(Year), df_nifes_cod2)   
```

## 9. Create df_median and set parameter names  
Note that we use Flag = "" for data over LOQ (not ">" as in the original script)  
```{r}
df_median <- df_nifes_cod2 %>% 
  group_by(Year, Uttaksområde, Organ, Parameter) %>%
  summarise(Conc = median(Conc, na.rm = TRUE), N = n(), Over_LOQ = sum(Flag %in% ""))
head(df_median, 3)

# Change parameter names (only those needed for indicators)
sel <- df_median$Parameter %in% "Cd"; sum(sel)
df_median$Parameter[sel] <- "CD"
sel <- df_median$Parameter %in% "Hg"; sum(sel)
df_median$Parameter[sel] <- "HG"
sel <- df_median$Parameter %in% "Pb"; sum(sel)
df_median$Parameter[sel] <- "PB"
sel <- df_median$Parameter %in% "HCB"; sum(sel)    # None
sel <- df_median$Parameter %in% "BDE6S"; sum(sel)  # OK, doesn't need to be changed
sel <- df_median$Parameter %in% "pp-DDE"; sum(sel)  # None 
df_median$Parameter[sel] <- "DDEPP"
sel <- df_median$Parameter %in% "Sum PCB 7"; sum(sel)   # 8
df_median$Parameter[sel] <- "CB_S7"

# HCB, pp-DDE
cat("Parameters:\n", table(df_median$Parameter) %>% names() %>% paste(collapse = "; "))
```

## 10. Save 
```{r}
# dir.create("Data")

if (save_data){
  saveRDS(df_nifes_cod2, "Data/01_df_nifes_cod2.rds")
  saveRDS(df_median, "Data/01_df_median.rds")
}

```


