---
title: "Get levels/trends for bluemussel, 2021"
output: html_document
---

- The file is called *2021* to keep in line with the other 2021 files (done in late 2020 for publishing in 2021)  
- Note the data is saved in a K folder called *2020*    
- Note that this uses data until *2019*  
- Resulting data is also copied to `K:\Avdeling\Mar\NOG\JMGdata\Kart\Indikatorer\2020\Grunnlagsdata` 


## Packages + functions  
```{r}

library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)
library(maps)
library(mapdata)

# File name for export
fn_export <- "Data_export/MytiEdu_Snegl_2020_ver3.csv"
overwrite <- FALSE

```

## Data
### Read main NIVA data  
Only like 10 seconds (3-4 minutes if column types is not given)  
__NOTE hard-coded ranges for excel sheet (both no. of columns and rows); 2 places__
```{r}
# Define column types for faster reading + fewer (or none) errors

# Column types - NOTE: for data going until 2019 
# Until column DT only
coltypes = c(rep("text",18),                  # PARAM - "Reference stations" [column A - R]
             rep("numeric", 4),               # "Reference station count" - "PROREF" [column S - V]
             rep(c("numeric","text"),40),     # Data 1980 - 2019 (2 columns per year) [W - CX] - increases by 2 each year
             "text","numeric","numeric","numeric","numeric",  # "N_string" - "EQS-threshold" last year [CY-DC]
             "text","numeric","numeric","numeric","numeric",  # "N_string" - "EQS-threshold" this year [DD-DH]
             "text",                                          # Dummy [DI]
             rep("numeric", 10),                              # "Trend p(long) this year" - "No. of Years(short) this year" [DJ - DS]
             "text"                           # "Trends this year" [DT]
             )

# write.csv2(coltypes, "clipboard", row.names = FALSE, col.names = FAL

# Old file
# fn <- "C:/Data/seksjon 212/Milkys_2018/Big_excel_table/Data_xl_lessthans_ver14_bluered.xlsm"

# New file
fn <- "C:/Data/seksjon 212/Milkys2_pc/Files_from_Jupyterhub_2019/Big_excel_table/Data_xl_2020-08-05_ver15.xlsm"
data_xl_orig <- read_excel(fn,
                      range = "A1:DT24347",    # NOTE HARD-CODED RANGE  
                      col_types = coltypes)

# Read the single less-than value 
data_xl_lessthans <- read_excel(fn,
                      range = "HM1:HM24347",    # NOTE HARD-CODED RANGE  
                      col_types = "logical")


# dim(data_xl)
# [1]  29546   216

```

### Stations
```{r}

# Station names + coordinates
folder_milkys <- "H:/Documents/seksjon 212/Milkys 2017/Analyse"
df_stations <- read_excel(paste0(folder_milkys, '/Stasjonsfiler/Milkys_koordinater_Sammenligning_NOG_0410_2017_forR.xlsx'), sheet = "for_R")
colnames(df_stations)[1:3] <- c("LATITUDE", "LONGITUDE", "STATION_CODENAME")
df_stations$STATION_CODE <- stringr::str_extract(df_stations$STATION_CODENAME, "([^[[:blank:]]]+)") 
df_stations$STATION_NAME <- stringr::str_extract(df_stations$STATION_CODENAME, "(?<=[[:blank:]]).*")
df_stations <- df_stations[,c("STATION_CODE", "STATION_NAME", "LONGITUDE", "LATITUDE")]
head(df_stations)

```

```{r}

data_xl <- data_xl_orig %>%
  rename(
    PARAM = `Parameter Code`,
    LATIN_NAME = Species,
    STATION_CODE = `Station Code`,
    Station.Name = `Station Name`,
    TISSUE_NAME = Tissue)

names(data_xl) <- sub("^V8", "Yr_198", names(data_xl))
names(data_xl) <- sub("^V9", "Yr_199", names(data_xl))
names(data_xl) <- sub("^V0", "Yr_200", names(data_xl))
names(data_xl) <- sub("^V1", "Yr_201", names(data_xl))

names(data_xl)

```
### Add VDSI for 36g Færder in 2019  
```{r}

sel <- with(data_xl, STATION_CODE == "36G" & grepl("VDSI", PARAM)); sum(sel)  

data_xl$Yr_2019[sel] <- 0

```


### Check if there is older data
```{r}

check <- data_xl %>% 
  filter(
    PARAM %in% c("CD", "HG", "PB", "BDE6S", "HCB", "DDEPP", "CB_S7") & 
    LATIN_NAME %in% "Mytilus edulis") %>%
  select(STATION_CODE, Station.Name, PARAM, Yr_2012, Yr_2013, Yr_2014, Yr_2015, Yr_2016, Yr_2017, Yr_2018, Yr_2019) %>%
  gather("Year", "Conc", Yr_2012:Yr_2019) %>%
  filter(!is.na(Conc))

# xtabs(~STATION_CODE + Year, check)
  
check <- left_join(check, df_stations, by = "STATION_CODE")
                   
check %>% 
  filter(LATITUDE > 62 & LONGITUDE < 22.2) %>%
  xtabs(~STATION_CODE + Year, .)

```

### 'data_xl_sel' = selected rows of excel table
```{r}

species <- c("Mytilus edulis", "Nucella lapillus", "N. lapillus / L. littorea")

# Parameters             
sel_param <- c("CD", "HG", "PB", "BDE6S", "HCB", "DDEPP", "CB_S7", "TBT", "VDSI", "VDSI/Intersex")

# Select records
sel_records <- with(data_xl, PARAM %in% sel_param & 
                             LATIN_NAME %in% species & 
                             Basis == "WW" & !is.na(Yr_2019))

sum(sel_records)   # 183
sum(data_xl$STATION_CODE == "71G")
sum(data_xl$STATION_CODE[sel_records] == "71G")

# Select rows
data_xl_sel <- data_xl[sel_records,]

# Add less-thans
data_xl_sel <- bind_cols(data_xl_sel, data_xl_lessthans[sel_records,])

# remove data_xl
rm(data_xl)

```





### Limits
```{r}
# Limits for EQS and Mattrygghet (food safety)
df_limits <- read_excel("Input_data/Grenseverdier_fra_Sylvia.xlsx")

```



## Make 'df_indicator'  
```{r}

df_indicator <- tibble(
  PROJECT_ID = 3699,
  LATIN_NAME = data_xl_sel[["LATIN_NAME"]],
  STATION_CODE = data_xl_sel[["STATION_CODE"]])

df_indicator <- left_join(df_indicator, df_stations, by = "STATION_CODE")
df_indicator$PARAM <- data_xl_sel[["PARAM"]]
df_indicator$N <- NA

# View(df_indicator)
nrow(df_indicator)
# 183

```


## Add Conc value for check  
To remove before export
```{r}

value_2019 <- data_xl_sel$Yr_2019

# Less-thans are set a tad lower, to be put in the lower class
sel2 <- !is.na(data_xl_sel$Lt_2019) & data_xl_sel$Lt_2019
value_2019[sel2] <- value_2019[sel2] - 0.00001

df_indicator$Conc <- value_2019

```

## Add SPECIES_ID, TISSUE_NAME, NIVA_CODE
```{r}
species_code <- 17

df_indicator$SPECIES_ID <- species_code

df_tissue_code <- tibble(
  TISSUE_NAME = c("Whole soft body", "Muskel", "Lever"),
  NIVA_CODE = c("SB","MU","LI")
  )

df_indicator$TISSUE_NAME <- data_xl_sel[["TISSUE_NAME"]]

n1 <- nrow(df_indicator)
df_indicator <- left_join(df_indicator, df_tissue_code, by = "TISSUE_NAME")
n2 <- nrow(df_indicator)

if (n2 != n1)
  stop("Number of rows changed - check whether 'df_tissue_code' has only unique values")


```

## Add trend
```{r}
# Check 'set_symbol' function: 
# source("18_Time_series_write_to_Excel_functions.R")
# set_symbol
df_symbol_to_trend <- tibble(
  trend_symbol = c("§", "«", "¢", "é", "ê"),
  trend_text = c("Too little data", "Too few data above LOQ", "No trend", "Trend up", "Trend down"),
  trend = c(0,0,1,2,3)
  )

#
# If Long-term trend:
# df_indicator$trend_symbol <- substr(data_xl_sel[sel_records, "Trends.2017"], 1, 1)   # 1st symbol, e.g. the "ê" in ê/¢
#
# If 10 year trend:
df_indicator$trend_symbol <- substr(data_xl_sel[["Trends this year"]], 3, 3)     # 3rd symbol, e.g. the "¢" in ê/¢
table(addNA(df_indicator$trend_symbol))

nrow(df_indicator)
df_indicator <- left_join(df_indicator, df_symbol_to_trend, by = "trend_symbol")
nrow(df_indicator)

# Check result
table(addNA(df_indicator$trend))

df_indicator$trend_symbol <- NULL
df_indicator$trend_text <- NULL

head(df_indicator, 3)

```
 
## Save (so far)  
```{r}
# Save indicator for NIVA data
saveRDS(df_indicator, "Data/05_df_indicator_bluemussel_2021.rds")
# df_indicator <- readRDS("Data/05_df_indicator_bluemussel.rds")
write.csv(df_indicator, "Data/05_df_indicator_bluemussel_2021.csv", quote = FALSE, row.names = FALSE)

# Save selected excel data
saveRDS(data_xl_sel, "Data/05_data_xl_sel_bluemussel_2021.rds")

```

## Map data
```{r, fig.width=4, fig.height=5}

# Check that all stations have positions
sum(is.na(df_indicator$LATITUDE))

df_indicator %>%
  count(STATION_CODE, STATION_NAME, LATITUDE,  LONGITUDE) %>%
  arrange(LATITUDE) %>%
  filter(LATITUDE >= 61.5)

plot(LATITUDE ~ LONGITUDE, df_indicator)
maps::map(regions = "Norway", add = TRUE, col = "brown")

```

## Filter df_indicator by position    
*Be sure to selet the right area*  
```{r}

# Norskehavet
# df_indicator <- df_indicator %>% 
#   filter(LATITUDE > 62 & LONGITUDE < 22.2)

# Nordsjøen
df_indicator <- df_indicator %>% 
  filter(LATITUDE < 62)

# 
tab <- df_indicator %>%
  count(STATION_NAME, LATITUDE,  LONGITUDE) %>%
  arrange(LONGITUDE) 

# Make STATION_NAME a factor with order = LONGITUDE
df_indicator <- df_indicator %>% 
  mutate(STATION_NAME = factor(STATION_NAME, levels = unique(tab$STATION_NAME)))

tab 


```
```{r}
# names(data_xl_sel)
```

## Proref, EQS and mattrygghet
```{r}

# Get PROREF (Q95) and EQS
df_q95 <- data_xl_sel[data_xl_sel$Basis %in% "WW", c("PARAM", "LATIN_NAME", "TISSUE_NAME", "PROREF", "EQS-threshold last year")] %>%
  group_by(PARAM, LATIN_NAME, TISSUE_NAME) %>%
  summarise(Q95 = first(`PROREF`), EQS_threshold = first(`EQS-threshold last year`),
            .groups = "drop")

# Check
# df_q95 %>% filter(PARAM == "HG")

# add PROREF and EQS thresholds
n1 <- nrow(df_indicator) # 94
df_indicator <- left_join(df_indicator, df_q95, 
                          by = c("LATIN_NAME", "PARAM", "TISSUE_NAME"))
n2 <- nrow(df_indicator) # 94

if (n2 != n1)
  stop("Number of rows changed - check whether df_q95 has only unique values")

# change EQS threshold for Hg (by mail from Norman; from 1.00 to 0.67)
sel <- df_indicator$PARAM %in% "CB_S7" 
df_indicator$EQS_threshold[sel] <- 0.67
cat("Changed EQS threshold for Hg -", sum(sel), "records")


# Classes changed - original classes were: c(-999999,1,2,5,10,20,999999)
df_indicator$KLASSE <- with(df_indicator, 
                             cut(Conc/Q95, breaks = c(-999999,1,2,10,999999), 
                                 right = FALSE, labels = FALSE)
                             )

cat("\n-------------------------------------------------\n")
cat("Tabulate PROREF class \n")
xtabs(~KLASSE + PARAM, df_indicator)
# KLASSE BDE6S CB_S7 CD HG PB
#      1     3     0  4  1  2
#      2     0     2  1  3  1
#      3     0     1  0  1  2


# EQS classe
df_indicator$EQS <- with(df_indicator, 
                          cut(Conc/EQS_threshold, breaks = c(-999999,1,999999), 
                              right = FALSE, labels = FALSE)
                          )

cat("\n-------------------------------------------------\n")
cat("Tabulate EQS class \n")
xtabs(~addNA(EQS) + PARAM, df_indicator)

# addNA(EQS) BDE6S CB_S7 CD HG PB
#       1        0     0  0  4  0
#       2        3     3  0  1  0
#       <NA>     0     0  5  0  5
      
```

## Add SPECIES_ID
```{r}

# We just set it to NA for simplicity :-)

df_indicator <- df_indicator %>%
  mutate(SPECIES_ID = case_when(
    LATIN_NAME %in% "Mytilus edulis" ~ "Blåskjell",
    LATIN_NAME %in% "Nucella lapillus" ~ "Purpursnegl",
    grepl("lapillus", LATIN_NAME) ~ "Purpursnegl/Strandsnegl")
  )
    

```

## Add food limit  
1 = below limit, 2 = above limit
```{r}

n1 <- nrow(df_indicator)
df_indicator <- left_join(
  df_indicator, 
  subset(df_limits, select = c(PARAM, LATIN_NAME, NIVA_CODE, Mattrygghet)), 
  by = c("PARAM", "LATIN_NAME", "NIVA_CODE"))
n2 <- nrow(df_indicator)

if (n2 != n1)
  stop("Number of rows changed - check whether 'df_limits' has only unique values")

# 96

df_indicator$Mattrygghet <- cut(with(df_indicator, Conc/Mattrygghet),
                                 breaks = c(-999999,1,999999), 
                                 right = FALSE, labels = FALSE)

xtabs(~addNA(Mattrygghet) + PARAM + TISSUE_NAME, df_indicator)

# addNA(Mattrygghet) BDE6S CB_S7 CD HG PB
#               1        0     3  5  0  5
#               <NA>     3     0  0  5  0
              
# tail(df_indicator)

```


## Save and export  
* In folder 'data': save RDS file with all variables   
* In folder 'Data_export': save csv file with selected variables (saved manually as Excel later)  
* File name for csv/excel file (`fn_export`) set in the start of the script
* The excel file is copied to `K:\Avdeling\Mar\NOG\JMGdata\Kart\Indikatorer\2020\Grunnlagsdata` 
```{r}

# Save with all variables
saveRDS(df_indicator, file = "Data/05_df_indicator_bluemussel_2021_ver1.RData")
# df_indicator <- readRDS( file = "Data/05_df_indicator_bluemussel_2021_ver1.RData")

# Save with selected variables
vars <- c("PROJECT_ID", "STATION_CODE", "STATION_NAME", 
  "LONGITUDE", "LATITUDE", "PARAM", "N", "KLASSE", 
  "SPECIES_ID", "trend", "NIVA_CODE", "EQS", "Mattrygghet")

# DOnt include HG in liver
sel <- with(df_indicator, !(PARAM == "HG" & TISSUE_NAME == "Lever"))

# setwd("H:/Documents/seksjon 212/Indikator 2018/Analyse")

#
# 
if (overwrite | !file.exists(fn_export)){
  write.csv2(df_indicator[sel,vars], file = fn_export, row.names = FALSE, na = "")
  cat("Data written to file", fn_export, "\n")
} else {
  warning("File", fn_export, "exists, data not saved.\nSet 'overwrite <- TRUE' to overwrite file.\n")
}


# check
if (FALSE){
  df1 <- read.csv2(file = "Data_export/MytiEdu_Snegl_2020_ver1.csv")
  df2 <- read.csv2(file = "Data_export/MytiEdu_Snegl_2020_ver2.csv")
  xtabs(~STATION_CODE, df1)
  xtabs(~STATION_CODE, df2)
}

```

## Checks
### Plot 1a
trend: 0 = no trend calculated, 1 = zero time trend, 2 = up, 3 = down  
```{r}
df <- df_indicator %>%
  mutate(Time_trend =
           case_when(trend %in% 0 ~ "Not calc.",
                     trend %in% 1 ~ "No trend",
                     trend %in% 2 ~ "Down",
                     trend %in% 3 ~ "Up")
         ) %>%
  mutate(Time_trend = factor(Time_trend, levels = c("Not calc.", "No trend", "Down", "Up")))

df %>%
  filter(PARAM %in% unique(PARAM)[1:3]) %>%
  ggplot(aes(STATION_NAME, Conc, fill = factor(Time_trend))) + 
  geom_col() + 
  scale_fill_manual(values = c("grey30", "orange", "green", "red")) +
  facet_wrap(~PARAM, scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

```



### Plot 1b
```{r, fig.width=9, fig.height=5}
df %>%
  filter(PARAM %in% unique(PARAM)[4:8]) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Time_trend)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("grey30", "orange", "green", "red")) +
  facet_wrap(~PARAM, scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))
```



### Plot 2 - trend maps     
```{r}

simple_map <- map_data("worldHires", c("Norway", "Sweden"))

df %>%
  ggplot(aes(LONGITUDE, LATITUDE, Conc, color = factor(Time_trend))) + 
  annotation_map(simple_map, fill = "lightgreen") +
  geom_point() +
  scale_color_manual(values = c("grey30", "orange", "green", "red")) +
  facet_wrap(~PARAM) +
  coord_map("lambert", parameters = c(10.4, 59.3))

```

### Plot 3 - mattrygghet
trend: 0 = no trend calculated, 1 = zero time trend, 2 = up, 3 = down  
```{r}

df %>%
  filter(!is.na(Mattrygghet)) %>%
  ggplot(aes(STATION_NAME, Conc, fill = factor(Mattrygghet))) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red")) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

```
### Plot 3b - maps     
```{r}

simple_map <- map_data("worldHires", c("Norway", "Sweden"))

df %>%
  ggplot(aes(LONGITUDE, LATITUDE, Conc, color = factor(Mattrygghet))) + 
  annotation_map(simple_map, fill = "grey60") +
  geom_point() +
  scale_color_manual(values = c("green", "red"), na.value = "grey20") +
  facet_wrap(~PARAM) +
  coord_map("lambert", parameters = c(10.4, 59.3))

```

