---
title: "Get NIVA data"
output: html_document
---



## Packages + functions  
```{r}
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)

# File name for export
fn <- "Data_export/MytiEdu_2018_ver1.csv"
overwrite <- FALSE

```

## Data
### Read main NIVA data  
Only like 10 seconds (3-4 minutes if column types is not given)  
__NOTE hard-coded ranges for excel sheet (both no. of columns and rows); 2 places__
```{r}
# Define column types for faster reading + fewer (or none) errors


# Column types, all columns
coltypes = c(rep("text",18),                  # PARAM - Backgr_stations
             rep("numeric", 4),               # Backgr_Nstat - Q95
             rep(c("numeric","text"),38),     # Data 1980 - 2017 (2 columns per year) [W - CS]
             rep(c("text","numeric","numeric","numeric","numeric"),2),  # Ant.prøver - EQS threshold, 2016+2017 [CV-DD]
             "text",                                          # OC 
             rep(c(rep("numeric",10), "text"),2),             # Trend p(long) - Trends, 2017+2016 [DF-EA]
             "text", rep("numeric",2), rep("text", 6),        # Last_two_years - Trends.2016.old [EB - EJ]
             rep(c("logical","text"),38),                     # Less-than columns (2 columns per year) [EK - HF]
             "text", "text"                                   # Trends.2017 V13 - V13 wrong [HG - HH]
             )

# Column types, until column DP only
coltypes = c(rep("text",18),                  # PARAM - Backgr_stations
             rep("numeric", 4),               # Backgr_Nstat - Q95
             rep(c("numeric","text"),38),     # Data 1980 - 2017 (2 columns per year) [W - CS]
             rep(c("text","numeric","numeric","numeric","numeric"),2),  # Ant.prøver - EQS threshold, 2016+2017 [CV-DD]
             "text",                                          # OC 
             rep("numeric",10), "text"                        # Trend p(long) - Trends, 2017 only [DF-DP]
             )

# write.csv2(coltypes, "clipboard", row.names = FALSE, col.names = FALSE)

fn <- "C:/Data/seksjon 212/Milkys_2018/Big_excel_table/Data_xl_lessthans_ver14_bluered.xlsm"
data_xl <- read_excel(fn,
                      range = "A1:DP29547",    # NOTE HARD-CODED RANGE  
                      col_types = coltypes)

# Read the single less-than value 
data_xl_lessthans <- read_excel(fn,
                      range = "HE1:HE29547",    # NOTE HARD-CODED RANGE  
                      col_types = "logical")


# dim(data_xl)
# [1]  29546   216

```
### Stations
```{r}
# Station names + coordinates
folder_milkys <- "H:/Documents/seksjon 212/Milkys 2017/Analyse"
df_stations <- read_excel(paste0(folder_milkys, '/Stasjonsfiler/Milkys_koordinater_Sammenligning_NOG_0410_2017_forR.xlsx'), sheet = "for_R")
colnames(df_stations)[1:3] <- c("LATITUDE", "LONGITUDE", "STATION_CODENAME")
df_stations$STATION_CODE <- stringr::str_extract(df_stations$STATION_CODENAME, "([^[[:blank:]]]+)") 
df_stations$STATION_NAME <- stringr::str_extract(df_stations$STATION_CODENAME, "(?<=[[:blank:]]).*")
df_stations <- df_stations[,c("STATION_CODE", "STATION_NAME", "LONGITUDE", "LATITUDE")]
head(df_stations)

```


### Check if there is older data
```{r}
check <- data_xl %>% 
  filter(
    PARAM %in% c("CD", "HG", "PB", "BDE6S", "HCB", "DDEPP", "CB_S7") & 
    LATIN_NAME %in% "Mytilus edulis") %>%
  select(STATION_CODE, Station.Name, PARAM, Yr_2010, Yr_2011, Yr_2012, Yr_2013, Yr_2014, Yr_2015, Yr_2016, Yr_2017) %>%
  gather("Year", "Conc", Yr_2010:Yr_2017) %>%
  filter(!is.na(Conc))

# xtabs(~STATION_CODE + Year, check)
  
check <- left_join(check, df_stations, by = "STATION_CODE")
                   
check %>% 
  filter(LATITUDE > 62 & LONGITUDE < 22.2) %>%
  xtabs(~STATION_CODE + Year, .)
```

### 'data_xl_sel' = selected rows of excel table
```{r}
species <- "Mytilus edulis"
#
# Muscle data (Hg)
#
sel_param <- c("CD", "HG", "PB", "BDE6S", "HCB", "DDEPP", "CB_S7")
sel_records <- with(data_xl, PARAM %in% sel_param & 
                             LATIN_NAME %in% species & 
                             Basis == "WW" & !is.na(Klasse.2017))

sum(sel_records)   # 183

# Select rows
data_xl_sel <- data_xl[sel_records,]

# Add less-thans
data_xl_sel <- bind_cols(data_xl_sel, data_xl_lessthans[sel_records,])

# remove data_xl
rm(data_xl)

```





### Limits
```{r}
# Limits for EQS and Mattrygghet (food safety)
df_limits <- read_excel("Input_data/Grenseverdier_fra_Sylvia.xlsx")

```



## Make 'df_indicator'  
```{r}
df_indicator <- tibble(
  PROJECT_ID = 3699,
  LATIN_NAME = data_xl_sel[["LATIN_NAME"]],
  STATION_CODE = data_xl_sel[["STATION_CODE"]])

df_indicator <- left_join(df_indicator, df_stations, by = "STATION_CODE")
df_indicator$PARAM <- data_xl_sel[["PARAM"]]
df_indicator$N <- NA

# View(df_indicator)
nrow(df_indicator)
# 183

```


## Add Conc value for check  
To remove before export
```{r}

value_2017 <- data_xl_sel$Yr_2017

# Less-thans are set a tad lower, to be put in the lower class
sel2 <- !is.na(data_xl_sel$Lt_2017) & data_xl_sel$Lt_2017
value_2017[sel2] <- value_2017[sel2] - 0.00001

df_indicator$Conc <- value_2017

```

## Add SPECIES_ID, TISSUE_NAME, NIVA_CODE
```{r}
species_code <- 17

df_indicator$SPECIES_ID <- species_code

df_tissue_code <- tibble(
  TISSUE_NAME = c("Whole soft body", "Muskel", "Lever"),
  NIVA_CODE = c("SB","MU","LI")
  )

df_indicator$TISSUE_NAME <- data_xl_sel[["TISSUE_NAME"]]

nrow(df_indicator)
df_indicator <- left_join(df_indicator, df_tissue_code, by = "TISSUE_NAME")
nrow(df_indicator)

```

## Add trend
```{r}
# Check 'set_symbol' function: 
# source("18_Time_series_write_to_Excel_functions.R")
# set_symbol
df_symbol_to_trend <- tibble(
  trend_symbol = c("§", "«", "¢", "é", "ê"),
  trend_text = c("Too little data", "Too few data above LOQ", "No trend", "Trend up", "Trend down"),
  trend = c(0,0,1,2,3)
  )

#
# If Long-term trend:
# df_indicator$trend_symbol <- substr(data_xl_sel[sel_records, "Trends.2017"], 1, 1)   # 1st symbol, e.g. the "ê" in ê/¢
#
# If 10 year trend:
df_indicator$trend_symbol <- substr(data_xl_sel[["Trends.2017"]], 3, 3)     # 3rd symbol, e.g. the "¢" in ê/¢
table(addNA(df_indicator$trend_symbol))

nrow(df_indicator)
df_indicator <- left_join(df_indicator, df_symbol_to_trend, by = "trend_symbol")
nrow(df_indicator)

# Check result
table(addNA(df_indicator$trend))

df_indicator$trend_symbol <- NULL
df_indicator$trend_text <- NULL

head(df_indicator, 3)

```
 
## Save (so far)  
```{r}
# Save indicator for NIVA data
saveRDS(df_indicator, "Data/05_df_indicator_bluemussel.rds")
# df_indicator <- readRDS("Data/05_df_indicator_bluemussel.rds")
write.csv(df_indicator, "Data/05_df_indicator_bluemussel.csv", quote = FALSE, row.names = FALSE)

# Save selected excel data
saveRDS(data_xl_sel, "Data/05_data_xl_sel_bluemussel.rds")

```

## Map data
```{r}
# Check that all stations have positions
sum(is.na(df_indicator$LATITUDE))

df_indicator %>%
  count(STATION_CODE, STATION_NAME, LATITUDE,  LONGITUDE) %>%
  arrange(LATITUDE) %>%
  filter(LATITUDE >= 61.5)

plot(LATITUDE ~ LONGITUDE, df_indicator)
maps::map(regions = "Norway", add = TRUE, col = "brown")
```

## Filter df_indicator by position
```{r}

df_indicator <- df_indicator %>% 
  filter(LATITUDE > 62 & LONGITUDE < 22.2)

tab <- df_indicator %>%
  count(STATION_NAME, LATITUDE,  LONGITUDE) %>%
  arrange(LATITUDE) 

# Make STATION_NAME a factor with order = latitude
df_indicator <- df_indicator %>% 
  mutate(STATION_NAME = factor(STATION_NAME, levels = tab$STATION_NAME))

tab 


```

## Proref, EQS and mattrygghet
```{r}
# Get PROREF (Q95) and EQS
df_q95 <- data_xl_sel[data_xl_sel$Basis %in% "WW", c("PARAM", "LATIN_NAME", "TISSUE_NAME", "Q95", "EQS")] %>%
  group_by(PARAM, LATIN_NAME, TISSUE_NAME) %>%
  summarise(Q95 = first(Q95), EQS_threshold = first(EQS))

# Check
# df_q95 %>% filter(PARAM == "HG")

# add PROREF and EQS thresholds
nrow(df_indicator) # 94
df_indicator <- left_join(df_indicator, df_q95)
nrow(df_indicator) # 94

# change EQS threshold for Hg (by mail from Norman; from 1.00 to 0.67)
sel <- df_indicator$PARAM %in% "CB_S7"; sum(sel) 
df_indicator$EQS_threshold[sel] <- 0.67

# Classes changed - original classes were: c(-999999,1,2,5,10,20,999999)
df_indicator$KLASSE <- with(df_indicator, 
                             cut(Conc/Q95, breaks = c(-999999,1,2,10,999999), 
                                 right = FALSE, labels = FALSE)
                             )

xtabs(~KLASSE + PARAM, df_indicator)
# KLASSE BDE6S CB_S7 CD HG PB
#      1     3     0  4  1  2
#      2     0     2  1  3  1
#      3     0     1  0  1  2


# EQS classe
df_indicator$EQS <- with(df_indicator, 
                          cut(Conc/EQS_threshold, breaks = c(-999999,1,999999), 
                              right = FALSE, labels = FALSE)
                          )

xtabs(~addNA(EQS) + PARAM, df_indicator)

# addNA(EQS) BDE6S CB_S7 CD HG PB
#       1        0     0  0  4  0
#       2        3     3  0  1  0
#       <NA>     0     0  5  0  5
      



```

## Add SPECIES_ID
```{r}

species_code <- 17
df_indicator$SPECIES_ID <- species_code

```

## Add food limit  
1 = below limit, 2 = above limit
```{r}

nrow(df_indicator)
df_indicator <- left_join(
  df_indicator, 
  subset(df_limits, select = c(PARAM, LATIN_NAME, NIVA_CODE, Mattrygghet)), 
  by = c("PARAM", "LATIN_NAME", "NIVA_CODE"))
nrow(df_indicator)
# 96

df_indicator$Mattrygghet <- cut(with(df_indicator, Conc/Mattrygghet),
                                 breaks = c(-999999,1,999999), 
                                 right = FALSE, labels = FALSE)

xtabs(~addNA(Mattrygghet) + PARAM + TISSUE_NAME, df_indicator)

# addNA(Mattrygghet) BDE6S CB_S7 CD HG PB
#               1        0     3  5  0  5
#               <NA>     3     0  0  5  0
              
# tail(df_indicator)

```

## Save and export  
* On 'data': save with all variables   
* On 'Data_export': pick selected variables  
```{r}
# Save with all variables
saveRDS(df_indicator, file = "Data/05_df_indicator_bluemussel_ver1.RData")

# Save with selected variables
vars <- c("PROJECT_ID", "STATION_CODE", "STATION_NAME", 
  "LONGITUDE", "LATITUDE", "PARAM", "N", "KLASSE", 
  "SPECIES_ID", "trend", "NIVA_CODE", "EQS", "Mattrygghet")

# DOnt include HG in liver
sel <- with(df_indicator, !(PARAM == "HG" & TISSUE_NAME == "Lever"))

# setwd("H:/Documents/seksjon 212/Indikator 2018/Analyse")

#
# 
if (overwrite | !file.exists(fn)){
  write.csv2(df_indicator[sel,vars], file = fn, row.names = FALSE, na = "")
  cat("Data written to file", fn, "\n")
} else {
  cat("File", fn, "exists, data not saved.\nSet 'overwrite <- TRUE' to overwrite file.\n")
}



# df_indicator <- read.csv2(file = "Data_export/GaduMor_2018_withNIFES_ver1.csv")

```

## Checks
### Plot 1
trend: 0 = no trend calculated, 1 = zero time trend, 2 = up, 3 = down  
```{r}
df <- df_indicator %>%
  mutate(Time_trend =
           case_when(trend %in% 0 ~ "Not calc.",
                     trend %in% 1 ~ "No trend",
                     trend %in% 2 ~ "Down",
                     trend %in% 3 ~ "Up")
         ) %>%
  mutate(Time_trend = factor(Time_trend, levels = c("Not calc.", "No trend", "Down", "Up")))

df %>%
  filter(PARAM %in% unique(PARAM)[1:3]) %>%
  ggplot(aes(STATION_NAME, Conc, fill = factor(Time_trend))) + 
  geom_col() + 
  scale_fill_manual(values = c("grey30", "orange", "green", "red")) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))
```

### Plot 2
```{r}
df %>%
  filter(PARAM %in% unique(PARAM)[4:7]) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Time_trend)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("grey30", "orange", "green", "red")) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))
```

### Plot 3 - mattrygghet
trend: 0 = no trend calculated, 1 = zero time trend, 2 = up, 3 = down  
```{r}
df %>%
  filter(!is.na(Mattrygghet)) %>%
  ggplot(aes(STATION_NAME, Conc, fill = factor(Mattrygghet))) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red")) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))
```

