---
title: "Import and fix HI/NIFES data for 2022 report"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
---

Based on code from '01_NIVA_data_with_NIFES_3.R' ("H:/Documents/seksjon 212/Indikator 2018/Analyse")  

## 1. Packages  
```{r setup, include=FALSE}

library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)
library(purrr)

source("11_Import_and_fix_NIFES_data_2021_functions.R")

save_data <- TRUE
# save_data <- FALSE

```

## 2. Data  
```{r}
# dir("Input_data")

fn <- "Input_data/Cd Pb torsk lever 2006-2022 HI.xlsx"
# excel_sheets(fn)
dat1 <- read_excel(fn, col_types = "text")

fn <- "Input_data/Kvikksølv torskefilet Bhav 2006-2022 HI.xlsx"
# excel_sheets(fn)
dat2 <- read_excel(fn, col_types = "text") %>%
  rename(Prøvenr. = `Prøve nr.`)      # fix this column name (equal to dat1)

fn <- "Input_data/POPs torskelever Bhav 2006-2022 inkl. PBDE HI.xlsx"
# excel_sheets(fn)
dat3 <- read_excel(fn, col_types = "text") %>%
  rename(Prøvenr. = `Prøve nr.`)      # fix this column name (equal to dat1)

```

### Check column names
```{r}
data.frame(
  n1 = colnames(dat1)[1:10],
  n2 = colnames(dat2)[1:10],
  n3 = colnames(dat3)[1:10]
)

data.frame(
  n1 = colnames(dat1)[11:14],
  n2 = colnames(dat2)[11:14],
  n3 = colnames(dat3)[11:14]
)
  
# cat("\n")
# colnames(dat2)
```


### Check sample identifier Prøvenr.     
```{r}
# All should be zero - not the case for dat2 (16)
for (dat in list(dat1, dat2, dat3))
  print(
    dat %>%
      add_count(Prøvenr.) %>%
      filter(n > 1) %>% nrow()
  )

# If we now run the stuff on top again, it's OK (zero for all 3)

```
#### Fix dat2  

```{r}

# Prøvenr. lacking for all in dat2:
dat2 %>%
  count(Jnr, Prøvenr.) %>%
  filter(n > 1)

# Set prøvenr. for dat2:
# dat2 <- dat2 %>%
#   ungroup() %>%
#   # group_by(Jnr) %>%
#   group_by(År, Posisjon) %>%
#   mutate(Prøvenr. = 1:length(År) %>% as.character())

dat2$Prøvenr. <- as.character(1:nrow(dat2))

```

#### Fix dat3  

- Prøvenr. lacking the part after the slash

```{r}

# Prøvenr. lacking for all in dat2:
check <- dat3 %>%
  select(Jnr, Prøvenr.) %>%
  add_count(Prøvenr.) %>%
  filter(n > 1)

affected_samples <- unique(check$Prøvenr.)

dat3 <- dat3 %>%
  mutate(
    Prøvenr. = case_when(
      Prøvenr. %in% affected_samples ~ Jnr,
      TRUE ~ Prøvenr.)
  )

```


### Check sample identifier Prøvenr.     
```{r}

check1 <- list(dat1, dat2, dat3) %>%
  map(
    ~ .x %>%
      select(Jnr, Prøvenr.) %>%
      add_count(Prøvenr.) %>%
      filter(n > 1)
  )

check2 <- map_int(check1, nrow)

if (sum(check2) > 0){
  stop(sum(sum(check2) > 0), " of the datasets have duplicate Prøvenr.! Check 'check1'")
}

```


## 3. Standardize and combine data  

### Check that there is one sample number 

```{r}

# Jnr, Prøvenr.
# for (dt in list(dat1, dat2, dat3)){
#   check <- dt %>%
#     count(Jnr, Prøvenr.) %>%
#     filter(n > 1)
#   print(nrow(check))
# }

# Only Prøvenr.
for (dt in list(dat1, dat2, dat3)){
  check <- dt %>%
    count(Prøvenr.) %>%
    filter(n > 1)
  print(nrow(check))
}

```


### Check sample identifier Prøvenr.     
```{r}

check1 <- list(dat1, dat2, dat3) %>%
  map(
    ~ .x %>%
      select(Jnr, Prøvenr.) %>%
      add_count(Prøvenr.) %>%
      filter(n > 1)
  )

check2 <- map_int(check1, nrow)

if (sum(check2) > 0){
  stop(sum(sum(check2) > 0), " of the datasets have duplicate Prøvenr.! Check 'check1'")
}

```


### Standardize and combine    
Put all data sets on long format and combine
```{r}

# Remove columns that we don't need  
dat1 <- dat1 %>%
  select(-c(`Engelsk beskrivelse`, `Latinsk navn`, `Produkt beskrivelse`))
dat2 <- dat2 %>%
  select(-c(`Engelsk beskrivelse`, `Latinsk navn`, `Produkt beskrivelse`))
dat3 <- dat3 %>%
  select(-c(`Engelsk beskrivelse`, `Latinsk navn`, `Produkt beskrivelse`))

if (FALSE){
  # Checking:
  nifes_standardize_data(dat1) %>% xtabs(~Parameter, .)
  nifes_standardize_data(dat2) %>% xtabs(~Parameter, .)
  nifes_standardize_data(dat3) %>% xtabs(~Parameter, .)
}

df_nifes_cod <- bind_rows(
  nifes_standardize_data(dat1),
  nifes_standardize_data(dat2),
  nifes_standardize_data(dat3)
)



```

### Check measurements where conversion failed  

```{r}

check <- df_nifes_cod %>% filter(is.na(Conc))

if (nrow(check) > 0)
  warning(nrow(check), " measurements were not converted to numeric values - look at 'check'")

# View(check)

```


### Check that there is one unique value per sample and parameter  

```{r}

check <- df_nifes_cod %>% 
  add_count(År, Posisjon, Organ, Prøvenr., Parameter) %>%
  filter(n > 1)

xtabs(~n, check)
xtabs(~År, check)
xtabs(~Parameter, check)

```

```{r}

#### testing nifes_standardize_data()   

if (FALSE){

  # debugonce(nifes_standardize_data)
# debugonce(nifes_standardize_conc)
test1 <- nifes_standardize_data(dat1)
test2 <- nifes_standardize_data(dat2)
test3 <- nifes_standardize_data(dat3)
table(test1$Parameter)
table(test2$Parameter)
table(test3$Parameter)

  
}



```


### Checking PBDEs   

```{r}

pars_bde6s <- c("PBDE 28","PBDE 47","PBDE 99","PBDE 100","PBDE 153","PBDE 154")

ggplot(df_nifes_cod %>% filter(Parameter %in% pars_bde6s), 
       aes(as.numeric(År), Conc, color = Flag)) + 
  geom_jitter(width = 0.2) + 
  facet_grid(vars(Parameter), vars(Område))

```


## 4. Missing values  

### Check missing values  
```{r}
# 1. Missing Conc values (list of parameters)
tab <- xtabs(~Parameter + is.na(Conc), df_nifes_cod)
tab[tab[,2] > 0,]
cat(sum(tab[,2]), "missing Conc values\n")

sel1 <- is.na(df_nifes_cod$Conc) & !is.na(df_nifes_cod$Conc_chr) 
cat(sum(sel1), "missing Conc values with non-missing values of Conc_chr\n")

sel2 <- df_nifes_cod$Comment != ""
cat(sum(sel2), "'starred' Conc_chr values\n")

# 2. Missing Conc but unmissing Conc_chr values
# All have "n.a."
df_nifes_cod[sel1,] %>% select(Jnr, Parameter, Conc_chr, Conc, Unit, Flag)

# 3. Starred values
df_nifes_cod[sel2,] %>% select(Jnr, Parameter, Conc_chr, Conc, Unit, Flag, Comment)

```

### Remove missing values  
```{r}

df_nifes_cod <- df_nifes_cod %>%
  filter(!is.na(Conc))

```


## 5. Add Lat, Long

### Check lacking positions  
```{r}
# Just three positions, and some (125) missing 
df_nifes_cod %>% 
  count(Posisjon)

# All are Hg in cod fillet
df_nifes_cod %>%
  mutate(Has_position = ifelse(is.na(Posisjon),"Lack_position", "Has_position")) %>%
  count(Art, Organ, Parameter, Has_position) %>%
  pivot_wider(names_from = Has_position, values_from = n) %>%
  filter(!is.na(Lack_position))

# One Lnr may have several positions
#  (So we cant use Jnr for fixing position)
df_nifes_cod %>% 
  filter(!is.na(Posisjon)) %>%
  count(Jnr, Posisjon) %>%
  count(Jnr) %>%
  rename(No_of_positions_per_Jnr = n) %>%
  count(No_of_positions_per_Jnr)

```

### Remove lacking positions  
```{r}

df_nifes_cod <- df_nifes_cod %>%
  filter(!is.na(Posisjon))

```

### Extract Lat,long
```{r}

df_nifes_cod <- df_nifes_cod %>%
  # Split spring
  separate(Posisjon, c("LATITUDE","LONGITUDE"), sep = " ", remove = FALSE) %>%
  # Remove letters
  mutate(LATITUDE = sub("[a-zA-Z]", "", LATITUDE),
         LONGITUDE = sub("[a-zA-Z]", "", LONGITUDE)) %>%
  # Get LATITUDE, LONGITUDE (first replace decimal comma by '.', then convert)
  mutate(LATITUDE = as.numeric(sub(",", ".", LATITUDE)),
         LONGITUDE = as.numeric(sub(",", ".", LONGITUDE))
  )

df_nifes_cod %>% 
  count(Posisjon, LATITUDE, LONGITUDE)

```

## 6. More fixing     

### Fix year     

Don't have date, so we can't set Measurement year = 1.april to 31. March...    
```{r}

# df_nifes_cod %>% 
#   count(År)

df_nifes_cod <- df_nifes_cod %>% 
  mutate(Year = as.numeric(År))

df_nifes_cod %>% 
  count(År)

```

### Fix tissue  

Both Filet and Muskel, set all to Muskel  

```{r}

# Both Filet and Muskel, 
# xtabs(~Year + Organ + Art, df_nifes_cod)  

df_nifes_cod <- df_nifes_cod %>% 
  mutate(Organ = ifelse(Organ %in% "Filet", "Muskel", Organ))

xtabs(~Year + Organ + Art, df_nifes_cod)  

```


###  Add BDE6S

#### Checking PBDEs   

```{r}

pars_bde6s <- c("PBDE 28","PBDE 47","PBDE 99","PBDE 100","PBDE 153","PBDE 154")

ggplot(df_nifes_cod %>% filter(Parameter %in% pars_bde6s), 
       aes(Year, Conc, color = Flag)) + 
  geom_jitter(width = 0.2) + 
  facet_grid(vars(Parameter), vars(Område))

```


#### Add sum
```{r}

df_nifes_cod_BACKUP <- df_nifes_cod
# If error, restore the original
# df_nifes_cod <- df_nifes_cod_BACKUP

pars_bde6s <- c("PBDE 28","PBDE 47","PBDE 99","PBDE 100","PBDE 153","PBDE 154")

table(addNA(df_nifes_cod$Flag))

df_nifes_cod_extra <- df_nifes_cod %>%
  filter(Parameter %in% pars_bde6s & !is.na(Flag)) %>%
  group_by(Year, Posisjon, LONGITUDE, LATITUDE, Art, Organ, Prøvenr.) %>%
  summarise(Conc = sum(Conc),
            Flag = ifelse(sum(Flag %in% "") > 1, "", "<"),
            N_congeners = n()
            ) %>%
  ungroup() %>%
  mutate(Parameter = "BDE6S")

xtabs(~Year + N_congeners, df_nifes_cod_extra)

```


#### Check sum value  
```{r}

# df_nifes_cod_extra

df_nifes_cod_extra %>%
  ggplot(aes(Year, Conc, color = Posisjon)) +
  geom_jitter(size = 1) +
  facet_wrap(vars(Posisjon))

# df_nifes_cod %>% 
#   filter(Prøvenr. == "2019-390" & Parameter %in% pars_bde6s) %>%
#   View("example1")

df_nifes_cod %>% 
  count(Year, Prøvenr., Parameter) %>%
  arrange(n, Year) %>%
  filter(n > 1)


```
#### Add BDE sum to the rest

```{r}

if (!"BDE6S" %in% unique(df_nifes_cod$Parameter)){
  df_nifes_cod <- bind_rows(
    df_nifes_cod,
    df_nifes_cod_extra
  )
}

```

#### Check individual BDE and the sum  
```{r}

df_nifes_cod %>%
  filter(Parameter %in% c(pars_bde6s, "BDE6S")) %>%
  ggplot(aes(Year, Conc, color = Posisjon)) +
  geom_jitter(size = 1) +
  facet_wrap(vars(Parameter))

df_nifes_cod %>%
  filter(Parameter %in% c(pars_bde6s, "BDE6S")) %>% 
  group_by(Parameter, Year, Posisjon, LONGITUDE, LATITUDE, Art, Organ) %>%
  summarise(Above_LOQ = mean(Flag %in% "")) %>%
  ggplot(aes(Year, Above_LOQ, color = Posisjon)) +
  geom_point() +
  facet_wrap(vars(Parameter))


```


## 7. Checks
```{r}

unique(df_nifes_cod$Parameter)

# unique(df_nifes_cod$Parameter) %>% dput()

```

### Test plot  
Annual medians of the most important parameters
```{r, fig.width=9, fig.height=7}

df_nifes_cod %>%
  filter(Parameter %in% c("Cd", "Pb", "Hg", "PCB-118", "Sum PCB 7", 
                          "Hexachlorobenzene (HCB)", "pp-DDE", "PBDE 47", "BDE6S")) %>%
  group_by(Year, Posisjon, Parameter, Organ) %>%
  mutate(Conc_median = median(Conc)) %>%
  ggplot(aes(Year, Conc_median, color = Posisjon)) +
  geom_smooth(method = "lm") + geom_point() +
  facet_wrap(vars(Parameter, Organ), scales = "free_y")

```

## 8. Put on same format as NIVA data  

### a. Add "station code"  

```{r}

df_nifes_stations <- df_nifes_cod %>%
  distinct(LONGITUDE, LATITUDE) %>%
  mutate(STATION_CODE = paste0("HI_Barents", 1:n()))

if (!"STATION_CODE" %in% names(df_nifes_cod)){
  df_nifes_cod <- df_nifes_cod %>%
    left_join(df_nifes_stations, by = c("LATITUDE", "LONGITUDE"))
}

```

### b. Set parameter names  
```{r}

df_nifes_cod <- df_nifes_cod %>%
  mutate(PARAM = case_when(
    Parameter %in% "Cd" ~ "CD",
    Parameter %in% "Pb" ~ "PB",
    Parameter %in% "Hg" ~ "HG",
    Parameter %in% "Sum PCB 6" ~ "CB_S6",
    Parameter %in% "Sum PCB 7" ~ "CB_S7",
    Parameter %in% "Hexachlorobenzene (HCB)" ~ "HCB",
    Parameter %in% "pp-DDE" ~ "DDEPP",     # the rest of the DDT compounds we ignore
    grepl("^PCB", Parameter) ~  sub("PCB-", "CB", Parameter, fixed = TRUE),
    grepl("^PBDE", Parameter) ~  sub("PBDE", "BDE", Parameter, fixed = TRUE),
    TRUE ~ Parameter)
    )

if (FALSE)
  count(df_nifes_cod, Parameter, PARAM)

```

### Save  
```{r}

if (save_data){
  saveRDS(df_nifes_cod, "Data/11_df_nifes_cod (2022).rds")
}

```


## 9. Create df_median and set parameter names  
Note that we use Flag = "" for data over LOQ (not ">" as in the original script)  
```{r}

# Create df_median
df_median <- df_nifes_cod %>% 
  group_by(Year, Posisjon, LONGITUDE, LATITUDE, Art, Organ, Parameter) %>%
  summarise(Conc = median(Conc), 
            Conc_min = min(Conc), Conc_max = max(Conc),
            N = n(), Over_LOQ = sum(Flag %in% "")
            ) %>%
  ungroup()


# set parameter names
df_median <- df_median %>%
  mutate(Parameter = sub("PCB-", "CB", Parameter, fixed = TRUE)) %>% # xtabs(~Parameter, .)
  mutate(Parameter = sub("PBDE ", "BDE", Parameter, fixed = TRUE)) %>% # xtabs(~Parameter, .)
  mutate(Parameter = case_when(
    Parameter %in% "Cd" ~ "CD",
    Parameter %in% "Pb" ~ "PB",
    Parameter %in% "Hg" ~ "HG",
    Parameter %in% "Sum PCB 7" ~ "CB_S7",
    Parameter %in% "Hexachlorobenzene (HCB)" ~ "HCB",
    Parameter %in% "pp-DDE" ~ "DDEPP",     # the rest of the DDT compounds we ignore
    TRUE ~ Parameter)
    )

unique(df_median$Parameter)


# Check all
# cat("Parameters:\n", table(df_median$Parameter) %>% names() %>% paste(collapse = "; "))
```


## 10. Save 
```{r}
# dir.create("Data")

if (save_data){
  saveRDS(df_nifes_cod, "Data/11_df_nifes_cod (2022).rds")
  saveRDS(df_median, "Data/11_df_median (2022).rds")
}

# df_median <- readRDS("Data/01_df_median.rds")
# xtabs(~Parameter + Organ, df_median)

```


