---
title: "Import and fix HI/NIFES data for 2020 report"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
---

Based on code from '11_Import_and_fix_NIFES_data_2020.Rmd' ("H:/Documents/seksjon 212/Indikator 2018/Analyse")   
* NOTE: no new cod were caught 2019-2020, but some pesticides are added fro existing samples  

## 1. Packages  
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)

source("11_Import_and_fix_NIFES_data_2020_functions.R", encoding = "UTF-8")

# save_data <- TRUE
save_data <- FALSE

```

## 2. Data    
* NOTE: we simply skip the "metals in muscle" sheet since we don't need it here   
```{r}
dir("Input_data")

fn <- "Input_data/Data kysttorsk NH nov. 2018 per 160421.xlsx"
# excel_sheets(fn)

dat1 <- read_excel(fn, col_types = "text", sheet = "Metaller og POPs i torskelever", skip = 2)

#
# We don't need more, as the only thing we will do is to check trend in pesticides 
#

```

### Check column names   
* NOTE: Columns are a bit different, there are some extra columns  
```{r}

dat1$Uttaksdato <- NULL
dat1$Mottaksdato <- NULL
# dat1$Uttakssted <- NULL
dat1$Oppdragsgiver <- NULL
dat1$`Prosjekt notat` <- NULL
dat1$`Prøve status` <- NULL
dat1$`Prosjekt Nr.` <- NULL

dat1 <- dat1 %>%
  rename(`Prøvenr.` = Merking)

colnames(dat1)[1:10]    # same as before
colnames(dat1)[11:14]  # start of actual data

# These are the 10 forst columns that we need:
# dput(colnames(dat1)[1:10])
# c("Jnr", "Prøvenr.", "År", "Posisjon", "Uttaksområde", "Art", 
#   "Organ", "Engelsk beskrivelse", "Latinsk navn", "Produkt beskrivelse")

ncol(dat1)

colnames(dat1)[seq(11, 165, by = 2)]
colnames(dat1)[seq(12, 164, by = 2)]  
seq(12, 164, by = 2)[62]

colnames(dat1)[seq(11, 131, by = 2)]
colnames(dat1)[seq(12, 132, by = 2)]  
colnames(dat1)[133]                       # CB118 - no attached unit
colnames(dat1)[seq(134, 156, by = 2)]   # PBDEs
colnames(dat1)[seq(135, 157, by = 2)]   
colnames(dat1)[seq(158, 165, by = 2)]   # HCB, DDT compumnds


```


### Check sample identifiers   
```{r}

dat1 %>%
  count(Jnr, Prøvenr.) %>%
  filter(n > 1)

```



## 3. Standardize and combine data
  
Put all data sets on long format and combine    
* Note that there are two PCB118 columns (one measured in 'TEQ pg/g ww ' and one without unit)   
    - Hee we donæt care, we are only looking fpr the pesticides anyway  
```{r}

if (FALSE){
  debugonce(nifes_standardize_conc)
  nifes_standardize_conc(dat1) %>% 
    filter(Jnr %in% "2016-2040/14") %>%
    select(Jnr, Prøvenr., Parameter, Conc_chr)
}

df_nifes_cod <- nifes_standardize_data(dat1) %>%
  filter(Parameter != "Underleverandør")     # last column

df_nifes_cod %>%
  filter(Jnr %in% "2016-2040/14") %>%
  select(Jnr, Prøvenr., Parameter, Conc, Unit)

# Original version:
# df_nifes_cod <- bind_rows(
#   nifes_standardize_data(dat1),
#   nifes_standardize_data(dat2),
#   nifes_standardize_data(dat3)
# )

```

## 4. Missing values  

### Check missing values  
```{r}

# 1. Missing Conc values (list of parameters)
tab <- xtabs(~Parameter + is.na(Conc), df_nifes_cod)
# tab[tab[,2] > 0,]    # show table
cat(sum(tab[,2]), "missing Conc values\n")
sel1 <- is.na(df_nifes_cod$Conc) & !is.na(df_nifes_cod$Conc_chr) 
cat(sum(sel1), "missing Conc values with non-missing values of Conc_chr\n")
sel2 <- df_nifes_cod$Comment != ""
cat(sum(sel2), "'starred' Conc_chr values\n")

# 2. Missing Conc but unmissing Conc_chr values
# All have "n.a."
if (sum(sel1) > 0){
  df_nifes_cod[sel1,] %>% select(Jnr, Parameter, Conc_chr, Conc, Unit, Flag)
}

# 3. Starred values
if (sum(sel2) > 0){
  df_nifes_cod[sel2,] %>% select(Jnr, Parameter, Conc_chr, Conc, Unit, Flag, Comment)
}

```

### Remove missing values  
```{r}

df_nifes_cod <- df_nifes_cod %>%
  filter(!is.na(Conc))

```


## 5. Add Lat, Long

### Check lacking positions    
* Note: Træna + Vikna will be joined to 'Helgeland'  
```{r}

# A bunch of positions, none missing 
df_nifes_cod %>% 
  count(Posisjon)

xtabs(~År + addNA(Uttaksområde), df_nifes_cod)

if (FALSE){
  # All are Hg in cod fillet
  df_nifes_cod %>%
    mutate(Has_position = ifelse(is.na(Posisjon),"Lack_position", "Has_position")) %>%
    count(Art, Organ, Parameter, Has_position) %>%
    pivot_wider(names_from = Has_position, values_from = n) %>%
    filter(!is.na(Lack_position))
  
  # One Lnr may have several positions
  #  (So we cant use Jnr for fixing position)
  df_nifes_cod %>% 
    filter(!is.na(Posisjon)) %>%
    count(Jnr, Posisjon) %>%
    count(Jnr) %>%
    rename(No_of_positions_per_Jnr = n) %>%
    count(No_of_positions_per_Jnr)
}

```

### Remove lacking positions  
```{r}

df_nifes_cod <- df_nifes_cod %>%
  filter(!is.na(Posisjon))

```

### Extract Lat,long   
* Not actually needed here, we will use 'Uttaksområde'    
```{r}

df_nifes_cod2 <- df_nifes_cod %>%
  # Split spring
  separate(Posisjon, c("LATITUDE","LONGITUDE"), sep = "\\s+", remove = FALSE) %>%
  # Remove letters
  mutate(LATITUDE = sub("[a-zA-Z]", "", LATITUDE),
         LONGITUDE = sub("[a-zA-Z]", "", LONGITUDE)) %>%
  # Get LATITUDE, LONGITUDE (first replace decimal comma by '.', then convert)
  mutate(LATITUDE = as.numeric(sub(",", ".", LATITUDE)),
         LONGITUDE = as.numeric(sub(",", ".", LONGITUDE))
  )

df_nifes_cod2 %>% 
  count(Posisjon, LATITUDE, LONGITUDE)

```

## 6. More fixing     

### Fix year     

Don't have date, so we can't set Measurement year = 1.april to 31. March...    
```{r}

# df_nifes_cod2 %>% 
#   count(År)

df_nifes_cod2 <- df_nifes_cod2 %>% 
  mutate(Year = as.numeric(År))

df_nifes_cod2 %>% 
  count(År)

```

### Fix tissue  

Both Filet and Muskel, set all to Muskel  

```{r}

# Both Filet and Muskel, 
# xtabs(~Year + Organ + Art, df_nifes_cod2)  

df_nifes_cod2 <- df_nifes_cod2 %>% 
  mutate(Organ = ifelse(Organ %in% "Filet", "Muskel", Organ))

xtabs(~Year + addNA(Organ) + Art, df_nifes_cod2)  

```


###  Add BDE6S

#### Add sum
```{r}

df_nifes_cod2_BACKUP <- df_nifes_cod2
# If error, restore the original
# df_nifes_cod2 <- df_nifes_cod2_BACKUP

pars_bde6s <- c("PBDE 28","PBDE 47","PBDE 99","PBDE 100","PBDE 153","PBDE 154")

df_nifes_cod2_extra <- df_nifes_cod2 %>%
  filter(Parameter %in% pars_bde6s) %>%
  group_by(Year, Posisjon, LONGITUDE, LATITUDE, Uttaksområde, Art, Organ, Prøvenr.) %>%   # include 'Uttaksområde' here
  summarise(Conc = sum(Conc),
            Flag = ifelse(sum(Flag %in% "") > 1, "", "<")
            ) %>%
  ungroup() %>%
  mutate(Parameter = "BDE6S")

if (!"BDE6S" %in% unique(df_nifes_cod2$Parameter)){
  df_nifes_cod2 <- bind_rows(
    df_nifes_cod2,
    df_nifes_cod2_extra
  )
}

```

#### Check individual BDE and the sum  
```{r}

df_nifes_cod2 %>%
  filter(Parameter %in% c(pars_bde6s, "BDE6S")) %>%
  ggplot(aes(Year, Conc, color = Uttaksområde)) +      # changed from Posisjon 
  geom_jitter(size = 1) +
  facet_wrap(vars(Parameter))

df_nifes_cod2 %>%
  filter(Parameter %in% c(pars_bde6s, "BDE6S")) %>% 
  group_by(Parameter, Year, Uttaksområde, LONGITUDE, LATITUDE, Art, Organ) %>%  # changed from Posisjon
  summarise(Above_LOQ = mean(Flag %in% "")) %>%
  ggplot(aes(Year, Above_LOQ, color = Uttaksområde)) +      # changed from Posisjon
  geom_point() +
  facet_wrap(vars(Parameter))


```


## 7. Checks
```{r}

unique(df_nifes_cod2$Parameter)

# unique(df_nifes_cod2$Parameter) %>% dput()

```

### Test plot   
Annual medians of the most important parameters   
* Note: HCB listed asd 'HCB', not 'Hexachlorobenzene (HCB)'  
* Note: pp-DDE listed as 'p,p'-DDE', not 'Hexachlorobenzene (HCB)'  
```{r, fig.width=9, fig.height=7}

df_nifes_cod2 %>%
  filter(Parameter %in% c("Cd", "Pb", "Hg", "PCB-118", "Sum PCB 7", 
                          "HCB", "p,p'-DDE", "PBDE 47", "BDE6S")) %>%
  group_by(Year, Uttaksområde, Parameter, Organ) %>%                         # changed from Posisjon 
  mutate(Conc_median = median(Conc)) %>%
  ggplot(aes(Year, Conc_median, color = Uttaksområde)) +
  geom_smooth(method = "lm") + geom_point() +
  facet_wrap(vars(Parameter, Organ), scales = "free_y")

```



## 9. Create df_median and set parameter names  
Note that we use Flag = "" for data over LOQ (not ">" as in the original script)  
```{r}

# Create df_median
df_median <- df_nifes_cod2 %>% 
  mutate(Posisjon = case_when(                            # NOTE: change definition of Posisjon here!
    Uttaksområde %in% c("Træna","Vikna") ~ "Helgeland",
    TRUE ~ Uttaksområde)
  ) %>%
  group_by(Year, Posisjon, Art, Organ, Parameter) %>%     # changed from Posisjon, Long, Lat
  summarise(Conc = median(Conc), 
            Conc_min = min(Conc), Conc_max = max(Conc),
            N = n(), Over_LOQ = sum(Flag %in% "")
            ) %>%
  ungroup()


# set parameter names
df_median <- df_median %>%
  mutate(Parameter = sub("PCB-", "CB", Parameter, fixed = TRUE)) %>% # xtabs(~Parameter, .)
  mutate(Parameter = sub("PBDE ", "BDE", Parameter, fixed = TRUE)) %>% # xtabs(~Parameter, .)
  mutate(Parameter = case_when(
    Parameter %in% "Cd" ~ "CD",
    Parameter %in% "Pb" ~ "PB",
    Parameter %in% "Hg" ~ "HG",
    Parameter %in% "Sum PCB 7" ~ "CB_S7",
    Parameter %in% "Hexachlorobenzene (HCB)" ~ "HCB",
    Parameter %in% "p,p'-DDE" ~ "DDEPP",     # the rest of the DDT compounds we ignore
    TRUE ~ Parameter)
    )

unique(df_median$Parameter)


# Check all
# cat("Parameters:\n", table(df_median$Parameter) %>% names() %>% paste(collapse = "; "))
```


## 10. Save 
```{r}
# dir.create("Data")

if (save_data){
  saveRDS(df_nifes_cod2, "Data/11_df_nifes_cod (Norwegian Sea 2020).rds")
  saveRDS(df_median, "Data/11_df_median (Norwegian Sea 2020).rds")
}

# df_median <- readRDS("Data/01_df_median.rds")
# xtabs(~Parameter + Organ, df_median)

```


