---
title: "Import and fix HI/NIFES data for 2020 report"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
---

Based on code from '01_NIVA_data_with_NIFES_3.R' ("H:/Documents/seksjon 212/Indikator 2018/Analyse")  

## 1. Packages  
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)

source("11_Import_and_fix_NIFES_data_2020_functions.R")

# save_data <- TRUE
save_data <- FALSE

```

## 2. Data  
```{r}
dir("Input_data")

fn <- "Input_data/Cd Pb torsk lever 2006-2019 HI.xlsx"
# excel_sheets(fn)
dat1 <- read_excel(fn, col_types = "text")

fn <- "Input_data/Kvikksølv torskefilet Bhav 2006-2019.xlsx"
# excel_sheets(fn)
dat2 <- read_excel(fn, col_types = "text") %>%
  rename(Prøvenr. = `Prøve nr.`)      # fix this column name (equal to dat1)

fn <- "Input_data/POPs torskelever Bhav 2006-2019 med PBDE.xlsx"
# excel_sheets(fn)
dat3 <- read_excel(fn, col_types = "text") %>%
  rename(Prøvenr. = `Prøve nr.`)      # fix this column name (equal to dat1)

```

### Check column names
```{r}
data.frame(
  n1 = colnames(dat1)[1:10],
  n2 = colnames(dat2)[1:10],
  n3 = colnames(dat3)[1:10]
)

data.frame(
  n1 = colnames(dat1)[11:14],
  n2 = colnames(dat2)[11:14],
  n3 = colnames(dat3)[11:14]
)
  
# cat("\n")
# colnames(dat2)
```


### Check sample identifiers   
```{r}
# All should be zero - not the case for dat2 (16)
for (dat in list(dat1, dat2, dat3))
  print(
    dat %>%
      count(Jnr, Prøvenr.) %>%
      filter(n > 1) %>% nrow()
  )

# Prøvenr. lacking for all in dat2:
dat2 %>%
  count(Jnr, Prøvenr.) %>%
  filter(n > 1)

# Set prøvenr. for dat2:
dat2 <- dat2 %>%
  group_by(Jnr) %>%
  mutate(Prøvenr. = 1:length(Jnr) %>% as.character())

# If we now run the stuff on top again, it's OK (zero for all 3)

```

## 3. Standardize and combine data
  
Put all data sets on long format and combine
```{r}

df_nifes_cod <- bind_rows(
  nifes_standardize_data(dat1),
  nifes_standardize_data(dat2),
  nifes_standardize_data(dat3)
)


```

## 4. Missing values  

### Check missing values  
```{r}
# 1. Missing Conc values (list of parameters)
tab <- xtabs(~Parameter + is.na(Conc), df_nifes_cod)
tab[tab[,2] > 0,]
cat(sum(tab[,2]), "missing Conc values\n")
sel1 <- is.na(df_nifes_cod$Conc) & !is.na(df_nifes_cod$Conc_chr) 
cat(sum(sel1), "missing Conc values with non-missing values of Conc_chr\n")
sel2 <- df_nifes_cod$Comment != ""
cat(sum(sel2), "'starred' Conc_chr values\n")

# 2. Missing Conc but unmissing Conc_chr values
# All have "n.a."
df_nifes_cod[sel1,] %>% select(Jnr, Parameter, Conc_chr, Conc, Unit, Flag)

# 3. Starred values
df_nifes_cod[sel2,] %>% select(Jnr, Parameter, Conc_chr, Conc, Unit, Flag, Comment)

```

### Remove missing values  
```{r}

df_nifes_cod <- df_nifes_cod %>%
  filter(!is.na(Conc))

```


## 5. Add Lat, Long

### Check lacking positions  
```{r}
# Just three positions, and some (125) missing 
df_nifes_cod %>% 
  count(Posisjon)

# All are Hg in cod fillet
df_nifes_cod %>%
  mutate(Has_position = ifelse(is.na(Posisjon),"Lack_position", "Has_position")) %>%
  count(Art, Organ, Parameter, Has_position) %>%
  pivot_wider(names_from = Has_position, values_from = n) %>%
  filter(!is.na(Lack_position))

# One Lnr may have several positions
#  (So we cant use Jnr for fixing position)
df_nifes_cod %>% 
  filter(!is.na(Posisjon)) %>%
  count(Jnr, Posisjon) %>%
  count(Jnr) %>%
  rename(No_of_positions_per_Jnr = n) %>%
  count(No_of_positions_per_Jnr)

```

### Remove lacking positions  
```{r}

df_nifes_cod <- df_nifes_cod %>%
  filter(!is.na(Posisjon))

```

### Extract Lat,long
```{r}

df_nifes_cod <- df_nifes_cod %>%
  # Split spring
  separate(Posisjon, c("LATITUDE","LONGITUDE"), sep = "[:blank:]", remove = FALSE) %>%
  # Remove letters
  mutate(LATITUDE = sub("[a-zA-Z]", "", LATITUDE),
         LONGITUDE = sub("[a-zA-Z]", "", LONGITUDE)) %>%
  # Get LATITUDE, LONGITUDE (first replace decimal comma by '.', then convert)
  mutate(LATITUDE = as.numeric(sub(",", ".", LATITUDE)),
         LONGITUDE = as.numeric(sub(",", ".", LONGITUDE))
  )

df_nifes_cod %>% 
  count(Posisjon, LATITUDE, LONGITUDE)

```

## 6. More fixing     

### Fix year     

Don't have date, so we can't set Measurement year = 1.april to 31. March...    
```{r}

# df_nifes_cod %>% 
#   count(År)

df_nifes_cod <- df_nifes_cod %>% 
  mutate(Year = as.numeric(År))

df_nifes_cod %>% 
  count(År)

```

### Fix tissue  

Both Filet and Muskel, set all to Muskel  

```{r}

# Both Filet and Muskel, 
# xtabs(~Year + Organ + Art, df_nifes_cod)  

df_nifes_cod <- df_nifes_cod %>% 
  mutate(Organ = ifelse(Organ %in% "Filet", "Muskel", Organ))

xtabs(~Year + Organ + Art, df_nifes_cod)  

```


###  Add BDE6S
```{r}

pars_bde6s <- c("PBDE 28","PBDE 47","PBDE 99","PBDE 100","PBDE 153","PBDE 154")

df_nifes_cod_extra <- df_nifes_cod %>%
  filter(Parameter %in% pars_bde6s) %>%
  group_by(Year, Posisjon, LONGITUDE, LATITUDE, Art, Organ) %>%
  summarise(Conc = sum(Conc)) %>%
  mutate(Parameter = "BDE6S")

if (!"BDE6S" %in% unique(df_nifes_cod$Parameter)){
  df_nifes_cod <- bind_rows(
    df_nifes_cod,
    df_nifes_cod_extra
  )
}

```


## 7. Checks
```{r}

unique(df_nifes_cod$Parameter)

# unique(df_nifes_cod$Parameter) %>% dput()

```

### Test plot  
Annual medians of the most important parameters
```{r, fig.width=9, fig.height=7}

df_nifes_cod %>%
  filter(Parameter %in% c("Cd", "Pb", "Hg", "PCB-118", "Sum PCB 7", 
                          "Hexachlorobenzene (HCB)", "pp-DDE", "PBDE 47", "BDE6S")) %>%
  group_by(Year, Posisjon, Parameter, Organ) %>%
  mutate(Conc_median = median(Conc)) %>%
  ggplot(aes(Year, Conc_median, color = Posisjon)) +
  geom_smooth(method = "lm") + geom_point() +
  facet_wrap(vars(Parameter, Organ), scales = "free_y")

```



## 9. Create df_median and set parameter names  
Note that we use Flag = "" for data over LOQ (not ">" as in the original script)  
```{r}

# Create df_median
df_median <- df_nifes_cod %>% 
  group_by(Year, Posisjon, LONGITUDE, LATITUDE, Art, Organ, Parameter) %>%
  summarise(Conc = median(Conc), 
            Conc_min = min(Conc), Conc_max = max(Conc),
            N = n(), Over_LOQ = sum(Flag %in% "")
            ) %>%
  ungroup()


# set parameter names
df_median <- df_median %>%
  mutate(Parameter = sub("PCB-", "CB", Parameter, fixed = TRUE)) %>% # xtabs(~Parameter, .)
  mutate(Parameter = sub("PBDE ", "BDE", Parameter, fixed = TRUE)) %>% # xtabs(~Parameter, .)
  mutate(Parameter = case_when(
    Parameter %in% "Cd" ~ "CD",
    Parameter %in% "Pb" ~ "PB",
    Parameter %in% "Hg" ~ "HG",
    Parameter %in% "Sum PCB 7" ~ "CB_S7",
    Parameter %in% "Hexachlorobenzene (HCB)" ~ "HCB",
    Parameter %in% "pp-DDE" ~ "DDEPP",     # the rest of the DDT compounds we ignore
    TRUE ~ Parameter)
    )

unique(df_median$Parameter)


# Check all
# cat("Parameters:\n", table(df_median$Parameter) %>% names() %>% paste(collapse = "; "))
```


## 10. Save 
```{r}
# dir.create("Data")

if (save_data){
  saveRDS(df_nifes_cod, "Data/11_df_nifes_cod (2020).rds")
  saveRDS(df_median, "Data/11_df_median (2020).rds")
}

# df_median <- readRDS("Data/01_df_median.rds")
# xtabs(~Parameter + Organ, df_median)

```


