---
title: "Import and fix HI/NIFES data for 2020 report"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
---

Based on code from '01_NIVA_data_with_NIFES_3.R' ("H:/Documents/seksjon 212/Indikator 2018/Analyse")   
  
**NOTE 'revised_data' parameter** (set to TRUE/FALSE)  
* Set to TRUE to use Sylvia's new data (on PCB) , sent in mail from Mroman 23.02.2021  
* Results are saved in files named 'rev 19.02.2021'  

## 1. Packages  
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)

source("11_Import_and_fix_NIFES_data_2021_functions.R", encoding = "UTF-8")

# save_data <- TRUE
save_data <- TRUE

```

## 2. Data  
```{r}
dir("Input_data")

revised_data <- TRUE
# revised_data <- FALSE

if (revised_data){
  fn <- "Input_data/Data torsk Nordsjøen 30.11.2020 til kart rev 19.02.2021.xlsx"
} else {
  fn <- "Input_data/Data torsk Nordsjøen 30.11.2020 til kart.xlsx"
}
# excel_sheets(fn)
dat1 <- read_excel(fn, sheet = "Cd og Pb i lever", col_types = "text") %>%
  rename(
    `Prøvenr.` = `Merking`)

dat2 <- read_excel(fn, sheet = "Metaller i filet", col_types = "text") %>%
  rename(
    `Prøvenr.` = `Merking`)
dat2 <- dat2[-c(3,4)]

dat3 <- read_excel(fn, sheet = "Org. miljøgifter i lever", col_types = "text") %>%
  rename(
    Jnr = `Jnr...1`,
    `Prøvenr.` = `Merking...2`)


```

### Check column names
```{r}

data.frame(
  n1 = colnames(dat1)[1:50],  # Cd and Pb: 15-18
  n2 = colnames(dat2)[1:50],  # Hg: 29-30
  n3 = colnames(dat3)[1:50]   # PCB-101 etc: 22-54 (conc. + unit) + 55-61 (conc. only)
)

# 

# colnames(dat3)

```

### Check locations   
```{r}

xtabs(~Uttaksområde , dat1)
xtabs(~Uttaksområde , dat2)
xtabs(~Uttaksområde , dat3)

if (FALSE){
  
  dat2 %>%
    count(Uttaksområde, Uttakssted)
  
  strsplit2 <- function(...){
    x <- strsplit(...)
    data.frame(c1 = x[[1]][1], c2 = x[[1]][2])
  }
  
  df <- dat2$Uttakssted %>% 
    purrr::map_dfr(strsplit2, split = " ")
  dat2$c1 <- df$c1
  dat2$c2 <- df$c2
  
  dat2 %>%
    count(Uttaksområde, Uttakssted, c1, c2)
  
  dat2 <- dat2 %>%
    mutate(
      Lat = as.numeric(c1),
      Lon = as.numeric(c2))
  
  dat2 %>%
    count(Uttaksområde, Uttakssted, Lat, Lon)
  
  dat2 <- dat2 %>%
    mutate(
      Lat2 = case_when(
        is.na(Lat) ~ as.numeric(substr(c1,1,2)) + as.numeric(substr(c1,3,4))/60,
        !is.na(Lat) ~ Lat),
      Lon2 = case_when(
        is.na(Lon) ~ as.numeric(substr(c2,1,3)) + as.numeric(substr(c2,4,5))/60,
        !is.na(Lon) ~ Lon)
    )
  
  dat2 %>%
    count(Uttakssted, c1, c2, Lat2, Lon2)
  
}


```


## 3. Standardize and combine data

```{r}

if (!revised_data){
  
  # PCB7 is twice:
  names(dat3)[c(36,39)]  # [1] "PCB7"      "Sum PCB 7"
  
  # We delete the first (which has no associated unit column)
  dat3$PCB7 <- NULL
  
} else {
  
  # PCB7 is twice:
  names(dat3)[36:38]  # [1] "PCB7"  "Sum PCB 6"  "Enhet...38" 
  
  # Add unit column to the right of PCB7 (we re-use col 38)
  # DO THIS ONLY ONCE!
  dat3 <- dat3[c(1:36, 38, 37:ncol(dat3))]
  
  # Check
  names(dat3)[36:39]  # [1] "PCB7" "Enhet...38" "Sum PCB 6"  "Enhet...38"

  
}

# BOTH VERSIONS
# 1. Cd and Pb: 15-18
# 2. Hg: 29-30
# 3. PCB-101 + PBDE: 22-53 (conc. + unit) + 54-60 (conc. only)

if (FALSE){
  # 1. Cd and Pb: 15-18
  names(dat1)[c(15,17)]   # "Cd" "Pb"

  # 2. Hg: 29-30
  names(dat2)[29]         # "Hg"

  # 3. PCB-101 etc: 22-54 (conc. + unit) + 55-61 (conc. only)
  names(dat3)[seq(22,53,by = 2)]   # "PCB-101"   "PCB-118"   "PCB-138" .... "PBDE 183"
  names(dat3)[seq(54,60,by = 1)]   # "HCB"    "op-DDD" "op-DDE" "op-DDT" "pp-DDD" "pp-DDE" "pp-DDT"

}



# debugonce(nifes_standardize_data)
df1 <- nifes_standardize_data(dat1[c(1:10, 15:18)]) 

df2 <- nifes_standardize_data(dat2[c(1:10, 29:30)]) %>%
  mutate(Art = "TORSK",
         Organ = "Filet")

df3a <- nifes_standardize_data(dat3[c(1:10, 22:54)])

rm(df3b)
df3b <- dat3[c(1:10, 55:60)] %>%
  pivot_longer(-(Jnr:Organ), names_to = "Parameter", values_to = "Conc_chr") %>%
  mutate(
      Conc = sub("<", "", Conc_chr, fixed = TRUE) %>% 
        sub("*", "", ., fixed = TRUE) %>%
        as.numeric(),
      # Find less-thans and make Flag variable
      Flag = ifelse(grepl("<", Conc_chr), "<", ""),
      Comment = ifelse(grepl("*", Conc_chr, fixed = TRUE), "*", "")
    )

df3 <- bind_rows(df3a, df3b)


```
  
Put all data sets on long format and combine
```{r}

df_nifes_cod <- bind_rows(df1, df2, df3) %>%
  filter(Parameter != "Kjønn")



```

## 4. Missing values  

### Check missing values  
```{r, results = 'hold'}
# 1a. Parameters with no missing values
cat("Parameters with no missing Conc values\n")
tab <- xtabs(~Parameter + is.na(Conc), df_nifes_cod)
tab[tab[,2] == 0,]

# 1b. Missing Conc values (list of parameters)
cat("\nParameters with some missing Conc values\n")
tab <- xtabs(~Parameter + is.na(Conc), df_nifes_cod)
tab[tab[,2] > 0,]
cat(sum(tab[,2]), "missing Conc values\n")
sel1 <- is.na(df_nifes_cod$Conc) & !is.na(df_nifes_cod$Conc_chr) 
cat(sum(sel1), "missing Conc values with non-missing values of Conc_chr\n")
sel2 <- df_nifes_cod$Comment != ""
cat(sum(sel2), "'starred' Conc_chr values\n")

if (FALSE){
# 2. Missing Conc but unmissing Conc_chr values
# All have "n.a."
df_nifes_cod[sel1,] %>% select(Jnr, Parameter, Conc_chr, Conc, Unit, Flag)

# 3. Starred values
df_nifes_cod[sel2,] %>% select(Jnr, Parameter, Conc_chr, Conc, Unit, Flag, Comment)
}

```

### Remove missing values  
```{r}

df_nifes_cod <- df_nifes_cod %>%
  filter(!is.na(Conc))

```


## 5. Add Lat, Long   

*We skip this this year*  



## 6. More fixing     

### Fix year     

Don't have date, so we can't set Measurement year = 1.april to 31. March...    
```{r}

xtabs(~addNA(År), df_nifes_cod)

df_nifes_cod <- df_nifes_cod %>% 
  mutate(Year = as.numeric(År))

```

### Fix tissue  

Both Filet and Muskel, set all to Muskel  

```{r}

# Both Filet and Muskel, 
# xtabs(~Year + Organ + Art, df_nifes_cod)  
xtabs(~addNA(Organ) + addNA(Art), df_nifes_cod)  

df_nifes_cod %>%
  filter(is.na(Art)) 

df_nifes_cod <- df_nifes_cod %>% 
  mutate(Organ = ifelse(Organ %in% "Filet", "Muskel", Organ))

xtabs(~Year + Organ + Art, df_nifes_cod)  

```


###  Add BDE6S

#### Add sum
```{r}

df_nifes_cod_BACKUP <- df_nifes_cod
# If error, restore the original
# df_nifes_cod <- df_nifes_cod_BACKUP

pars_bde6s <- c("PBDE 28","PBDE 47","PBDE 99","PBDE 100","PBDE 153","PBDE 154")

df_nifes_cod_extra <- df_nifes_cod %>%
  filter(Parameter %in% pars_bde6s) %>% # xtabs(~Parameter, .)
  # group_by(Year, Posisjon, LONGITUDE, LATITUDE, Art, Organ, Prøvenr.) %>%
  group_by(Year, Art, Organ, Jnr, Prøvenr., Uttaksområde) %>%
  summarise(Conc = sum(Conc),
            Flag = ifelse(sum(Flag %in% "") > 1, "", "<"),
            n_param = n(),
            .groups = "drop"
            ) %>%
  mutate(Parameter = "BDE6S")

# nrow(df_nifes_cod_extra)

if (!"BDE6S" %in% unique(df_nifes_cod$Parameter)){
  df_nifes_cod <- bind_rows(
    df_nifes_cod,
    df_nifes_cod_extra
  )
  cat(nrow(df_nifes_cod_extra), "rows added \n")
}

```

#### Check individual BDE and the sum  
```{r}

df_nifes_cod %>%
  filter(Parameter %in% c(pars_bde6s, "BDE6S")) %>%
  ggplot(aes(Year, Conc, color = Uttaksområde)) +
  geom_jitter(size = 1) +
  facet_wrap(vars(Parameter))

df_nifes_cod %>%
  filter(Parameter %in% c(pars_bde6s, "BDE6S")) %>% 
  # group_by(Parameter, Year, Posisjon, LONGITUDE, LATITUDE, Art, Organ) %>%
  group_by(Parameter, Uttaksområde, Year, Art, Organ) %>%
  summarise(Above_LOQ = mean(Flag %in% ""), .groups = "drop") %>%
  ggplot(aes(Year, Above_LOQ, color = Uttaksområde)) +
  geom_point() +
  facet_wrap(vars(Parameter))


```


## 7. Checks
```{r}

unique(df_nifes_cod$Parameter)

# unique(df_nifes_cod$Parameter) %>% dput()

```

### Test plot  
Annual medians of the most important parameters
```{r, fig.width=9, fig.height=7}

df_nifes_cod %>%
  filter(Parameter %in% c("Cd", "Pb", "Hg", "PCB-118", "Sum PCB 7", 
                          "Hexachlorobenzene (HCB)", "pp-DDE", "PBDE 47", "BDE6S")) %>%
  group_by(Year, Uttaksområde, Parameter, Organ) %>%
  mutate(Conc_median = median(Conc)) %>%
  ggplot(aes(Year, Conc_median, color = Uttaksområde)) +
  geom_smooth(method = "lm") + geom_point() +
  facet_wrap(vars(Parameter, Organ), scales = "free_y")

```



## 9. Create df_median and set parameter names  
Note that we use Flag = "" for data over LOQ (not ">" as in the original script)  
```{r}

# Create df_median
df_median <- df_nifes_cod %>% 
  group_by(Year, Art, Organ, Parameter) %>%
  summarise(Conc = median(Conc), 
            Conc_min = min(Conc), Conc_max = max(Conc),
            N = n(), Over_LOQ = sum(Flag %in% "")
            ) %>%
  ungroup()


# set parameter names
df_median <- df_median %>%
  mutate(Parameter = sub("PCB-", "CB", Parameter, fixed = TRUE)) %>% # xtabs(~Parameter, .)
  mutate(Parameter = sub("PBDE ", "BDE", Parameter, fixed = TRUE)) %>% # xtabs(~Parameter, .)
  mutate(Parameter = case_when(
    Parameter %in% "Cd" ~ "CD",
    Parameter %in% "Pb" ~ "PB",
    Parameter %in% "Hg" ~ "HG",
    Parameter %in% "Sum PCB 7" ~ "CB_S7",
    Parameter %in% "Hexachlorobenzene (HCB)" ~ "HCB",
    Parameter %in% "pp-DDE" ~ "DDEPP",     # the rest of the DDT compounds we ignore
    TRUE ~ Parameter)
    )

unique(df_median$Parameter)


# Check all
# cat("Parameters:\n", table(df_median$Parameter) %>% names() %>% paste(collapse = "; "))
```


## 10. Save 
```{r}
# dir.create("Data")

if (save_data){
  if (revised_data){
    saveRDS(df_nifes_cod, "Data/11_df_nifes_cod (2021) rev 19.02.2021.rds")
    saveRDS(df_median, "Data/11_df_median (2021) rev 19.02.2021.rds")
  } else {
    saveRDS(df_nifes_cod, "Data/11_df_nifes_cod (2021).rds")
    saveRDS(df_median, "Data/11_df_median (2021).rds")
  }
}

# df_median <- readRDS("Data/01_df_median.rds")
# xtabs(~Parameter + Organ, df_median)

```


