---
title: "Get NIVA trends, 2021 (data unitl 2020)"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    
---

Actually the script should be called get NIVA _trends_ because that is what we actually do   

NOTE:  
* This script is used for both '2021 North Sea' and '2021 Norweegian Sea' - set 'area' to the right ocean!    


## Packages + functions  
```{r, results='hide', message=FALSE}
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)

```

### Year  
```{r}

year_2020_data <- TRUE

```



### Area
```{r}

# area <- "North_Sea"
area <- "Norwegian_Sea"

```


## Data

### Read main NIVA data  

- Only like 10 seconds (3-4 minutes if column types is not given)  
- __NOTE hard-coded number of columns for excel sheet (4 places)__  
- Row nmber 'automatic'

```{r}
# Define column types for faster reading + fewer (or none) errors

if (year_2020_data){
  
  # FOR 2020 DATA
  # Column types - NOTE: for data going until 2019 
  # Until column DT only
  coltypes = c(rep("text",18),                  # PARAM - "Reference stations" [column A - R]
               rep("numeric", 4),               # "Reference station count" - "PROREF" [column S - V]
               rep(c("numeric","text"),41),     # Data 1980 - 2019 (2 columns per year) - INCREASE THIS NUMBER BY 1 EACH YEAR
               "text","numeric","numeric","numeric","numeric",  # "N_string" - "EQS-threshold" last year 
               "text","numeric","numeric","numeric","numeric",  # "N_string" - "EQS-threshold" this year 
               "text",                                          # Dummy [DI]
               rep("numeric", 10),                              # "Trend p(long) this year" - "No. of Years(short) this year" 
               "text"                           # "Trends this year" [DV]
  )
  
} else {
  
  # FOR 2019 DATA
  # Column types - NOTE: for data going until 2019 
  # Until column DT only
  coltypes = c(rep("text",18),                  # PARAM - "Reference stations" [column A - R]
               rep("numeric", 4),               # "Reference station count" - "PROREF" [column S - V]
               rep(c("numeric","text"),40),     # Data 1980 - 2019 (2 columns per year) [W - CX] - INCREASE THIS NUMBER BY 1 EACH YEAR
               "text","numeric","numeric","numeric","numeric",  # "N_string" - "EQS-threshold" last year [CY-DC]
               "text","numeric","numeric","numeric","numeric",  # "N_string" - "EQS-threshold" this year [DD-DH]
               "text",                                          # Dummy [DI]
               rep("numeric", 10),                              # "Trend p(long) this year" - "No. of Years(short) this year" [DJ - DS]
               "text"                           # "Trends this year" [DT]
  )
}


if (year_2020_data){
  
  #
  # FOR 2020
  #
  
  fn <- "../Milkys2_pc/Files_from_Jupyterhub_2020/Big_excel_table/Data_xl_2021-09-15_ver03.xlsm"
  
  # For getting number of rows:
  data_xl_firstcol <- read_excel(fn, range = cell_cols("A:B"))

  # 
  # HARD-CODED for 2020! last column of 'data_xl_orig' - increases by 2 for each year
  #
  last_column <- "DV"
  range_2020 <- paste0("A1:", last_column, nrow(data_xl_firstcol))
  
  data_xl_orig <- read_excel(fn,
                      range = range_2020,   
                      col_types = coltypes)

  # 
  # HARD-CODED for 2020! last column of 'data_xl_lessthans' - increases by 4 for each year
  #
  last_column <- "HQ"   
  range_2020_lessthan <- paste0(last_column, "1:", last_column, nrow(data_xl_firstcol))

  # Read the single less-than value 
  data_xl_lessthans <- read_excel(fn,
                      range = range_2020_lessthan,    
                      col_types = "logical")


  } else {
    
    #
    # FOR 2019
    #
    
    fn <- "C:/Data/seksjon 212/Milkys_2018/Big_excel_table/Data_xl_lessthans_ver14_bluered.xlsm"
    
    data_xl_orig <- read_excel(fn,
                               range = "A1:DT24347",    # NOTE HARD-CODED RANGE  
                               col_types = coltypes)
    
    # Read the single less-than value 
    data_xl_lessthans <- read_excel(fn,
                                    range = "HM1:HM24347",    # NOTE HARD-CODED RANGE  
                                    col_types = "logical")
    
    
}

data_xl <- data_xl_orig

# dim(data_xl)
# [1]  29546   216

```


### 'data_xl_sel' = selected rows of excel table
```{r}

species <- "Gadus morhua"
#
# Muscle data (Hg)
#
tissue <- "Muscle"
sel_param <- c("HG")
sel_records1 <- with(data_xl, `Parameter Code` %in% sel_param & 
                              Tissue %in% tissue & 
                              Species %in% species & 
                              Basis == "WW" & !is.na(`EQS-class this year`))

#
# Liver data (the rest)
#
tissue <- "Liver"
sel_param <- c("CD", "HG", "PB", "BDE6S", "HCB", "DDEPP", "CB_S7")
sel_records2 <- with(data_xl, `Parameter Code` %in% sel_param & 
                              Tissue %in% tissue & 
                              Species %in% species & 
                              Basis == "WW") # & !is.na(`EQS-class this year`))


#
# Blue mussel
#
species <- "Mytilus edulis"
tissue <- "Whole soft body"
sel_param <- c("HG", "CD", "HG", "PB", "BDE6S", "HCB", "DDEPP", "CB_S7")
sel_records3 <- with(data_xl, `Parameter Code` %in% sel_param & 
                              Tissue %in% tissue & 
                              Species %in% species & 
                              Basis == "WW" & !is.na(`EQS-class this year`))

# Combine muscel and liver data
sel_records <- sel_records1 | sel_records2 | sel_records3
sum(sel_records1)  # 17
sum(sel_records2)  # 75
sum(sel_records3)  # 75
sum(sel_records)   # 92

# Select rows
data_xl_sel <- data_xl[sel_records,]

# Add less-thans
data_xl_sel <- bind_cols(data_xl_sel, data_xl_lessthans[sel_records,])

# remove data_xl
# rm(data_xl)

xtabs(~`Parameter Code` + Tissue, data_xl_sel )

data_xl[sel_records2,] %>% xtabs(~`Parameter Code` + Tissue, . )

```


### Station data  
Shortened from script 98 in 'Milkys'   
```{r}

data_stations <- readxl::read_excel("../Milkys2_pc/Files_to_Jupyterhub_2019/Kartbase_edit.xlsx") %>%
  rename(LATITUDE = Lat, LONGITUDE = Long, 
         STATION_NAME = Station_name) %>%
  filter(!is.na(STATION_CODE))# %>%
  # One duplicate (different names), we just remove one of them
  # filter(!STATION_NAME %in% "Risøy, Østerfjord")

### Pick by area and species    

if (area == "North_Sea"){
  
  data_stations <- data_stations %>% 
    filter(LATITUDE < 61) %>%
    # no blue mussel, flatfish or snail stations
    filter(!(grepl("F", STATION_CODE) | grepl("G", STATION_CODE)))
  
} else if (area == "Norwegian_Sea"){
  
  data_stations <- data_stations %>% 
    filter(LATITUDE > 62.2 & LATITUDE < 70 & LONGITUDE < 22) %>%
    # no blue mussel, flatfish or snail stations
    filter(grepl("B", STATION_CODE))
  
}

data_stations

```


### Limits
```{r}
# Limits for EQS and Mattrygghet (food safety)
df_limits <- read_excel("Input_data/Grenseverdier_fra_Sylvia.xlsx")

```


## Add Conc value for check  
- Concnetration value of last year  
- If less-than, it is set slightly lower than value (in case the limit is exacly the same)
- To remove before export
```{r}

if (year_2020_data){
  
  value_lastyear <- data_xl_sel$V20          # HARD_CODED YEAR
  
  # Less-thans are set a tad lower, to be put in the lower class
  sel2 <- !is.na(data_xl_sel$Lt_2019) & data_xl_sel$Lt_2019
  
} else {
  
  value_lastyear <- data_xl_sel$V19          # HARD_CODED YEAR
  
  # Less-thans are set a tad lower, to be put in the lower class
  sel2 <- !is.na(data_xl_sel$Lt_2019) & data_xl_sel$Lt_2019
  
}

value_lastyear[sel2] <- value_lastyear[sel2] - 0.00001

data_xl_sel$Conc_for_check <- value_lastyear

```
### For checking  

```{r}

if (FALSE){
  
  data_xl %>% 
    filter(`Parameter Code` == "HG" & Basis == "WW" & `Station Code` == "43B2") %>%
    select(`Parameter Code`, Basis, `Station Code`, `Station Name`, PROREF, V19, V20)
  
  # Don't need Basis == "WW" - taht has been selected already
  data_xl_sel %>%
    filter(`Parameter Code` == "HG" & `Station Code` == "43B2") %>%
    select(`Parameter Code`, Basis, `Station Code`, `Station Name`, PROREF, V19, V20, Conc_for_check)
  
}

```


## 'df_indicator'  

### Make 'df_indicator'  
```{r}

df_indicator <- tibble(
  PROJECT_ID = 3699,
  LATIN_NAME = data_xl_sel[["Species"]],
  STATION_CODE = data_xl_sel[["Station Code"]],
  SPECIES_ID = 17,                  # HARD-CODED - check above that we only have cod stations
  TISSUE_NAME = data_xl_sel[["Tissue"]],
  PARAM = data_xl_sel[["Parameter Code"]],
  Conc = data_xl_sel[["Conc_for_check"]],
  trend_symbol = substr(data_xl_sel$`Trends this year`, 3, 3),     # 3rd symbol, e.g. the "¢" in ê/¢ - FOR 10-YEAR TREND
  N = NA
  )

#
# Add station position + name
#
nrow(df_indicator)  # 275
df_indicator <- df_indicator %>% 
  inner_join(data_stations, by = "STATION_CODE")
nrow(df_indicator)  # 44


```


### Add NIVA_CODE
```{r}

df_tissue_code <- tibble(
  TISSUE_NAME = c("Whole soft body", "Muscle", "Liver"),
  NIVA_CODE = c("SB","MU","LI")
  )

nrow(df_indicator)
df_indicator <- left_join(df_indicator, df_tissue_code, by = "TISSUE_NAME")
nrow(df_indicator)

```

### Add trend text
```{r}

# Check 'set_symbol' function: 
# source("18_Time_series_write_to_Excel_functions.R")
# set_symbol
df_symbol_to_trend <- tibble(
  trend_symbol = c("§", "«", "¢", "é", "ê"),
  trend_text = c("Too little data", "Too few data above LOQ", "No trend", "Trend up", "Trend down"),
  trend = c(0,0,1,2,3)
  )

# Add trend symbol  
nrow(df_indicator)
df_indicator <- left_join(df_indicator, df_symbol_to_trend, by = "trend_symbol")
nrow(df_indicator)

# Check result
table(addNA(df_indicator$trend))

df_indicator$trend_symbol <- NULL
df_indicator$trend_text <- NULL

head(df_indicator, 3)

```

## Save  
```{r}

overwrite <- FALSE
# overwrite <- TRUE

if (overwrite){
  
  if (area == "North_Sea"){
    
    # Save indicator for NIVA data
    saveRDS(df_indicator, "Data/13_df_indicator_NIVA_only (2021).rds")
    write.csv(df_indicator, "Data/13_df_indicator_NIVA_only (2021).csv", 
              quote = FALSE, row.names = FALSE)
    
    # Save selected excel data
    saveRDS(data_xl_sel, "Data/13_data_xl_sel (2021).rds")
    
    
  } else if (area == "Norwegian_Sea"){
    
    # Save indicator for NIVA data
    saveRDS(df_indicator, "Data/13_df_indicator_NIVA_only (2021_NorwSea).rds")
    write.csv(df_indicator, "Data/13_df_indicator_NIVA_only (2021_NorwSea).csv", 
              quote = FALSE, row.names = FALSE)
    
    # Save selected excel data
    saveRDS(data_xl_sel, "Data/13_data_xl_sel (2021_NorwSea).rds")
    
  }
  
}

if (FALSE) {
  
  df_indicator <- readRDS("Data/13_df_indicator_NIVA_only (2021_NorwSea).rds")

  
}

```

