---
title: "Combine NIVA and NIFES data 2022 - Norwegian Sea"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
---

Combining medians from   
* NIFES data (script 11) - special script for 2022_NorwSea
* Trends from NIFES data (script 12) - special script for 2022_NorwSea   
* Medians and trends from NIVA (script 13) - same script as 2022 North Sea (set 'area' to "Norweegian Sea')  
* In addition to merging the results, limit values are also added in this script    
  
The resulting summary data (one line per time series, i.e. station/substance) are saved as csv and excel files. These are also copied to `K:\Avdeling\Mar\NOG\JMGdata\Kart\Indikatorer\2020\Grunnlagsdata Norskehavet`      



## Packages + functions  
```{r}

library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)


# File name for export
fn <- "Data_export/GaduMor_2022data_ver01"
fn_csv <- paste0(fn, ".csv")
fn_xlsx <- paste0(fn, ".xlsx")
overwrite <- FALSE
#overwrite <- TRUE

# Background map
library(mapdata)
simple_map <- map_data("worldHires", "Norway") %>% fortify()

```

## Data
Indicators from NIVA serves as the starting point for combining the data  
```{r}

# NIVA medians
data_xl_sel <- readRDS("Data/13_data_xl_sel (2022).rds")

# NIVA indicators
df_indicator <- readRDS("Data/13_df_indicator_NIVA_only (2022).rds")

# NIFES medians
df_median <- readRDS("Data/11_df_median (2022).rds") %>%
  as.data.frame()

# NIFES regression
nifes_regression <- readRDS("Data/12_nifes_regression (2022).rds")

# Food and EQS limits
df_limits <- read_excel("Input_data/Grenseverdier_fra_Sylvia.xlsx")

# Parameters
sel_param <- c("CD", "HG", "PB", "HCB", "DDEPP", "CB_S7", "BDE6S")

```

### For 2021 specifically  
```{r}

df_2017 <- read_excel(
  "K:/Avdeling/Mar/NOG/JMGdata/Kart/Indikatorer/2017/Grunnlagsdata/GaduMor_2018_withNIFES_ver4.xlsx") 

df_2017_northsea <- df_2017 %>%
  filter(STATION_NAME == "Nordsjøen") %>%
  summarise(
    LONGITUDE = mean(LONGITUDE),
    LATITUDE = mean(LATITUDE)
    )

df_median <- df_median %>%
  mutate(
    Posisjon = "Nordsjøen",
    LONGITUDE = df_2017_northsea$LONGITUDE,
    LATITUDE = df_2017_northsea$LATITUDE)


```


## Make 'df_nifes_finalyear'   
```{r}

df_median <- df_median %>%
  mutate(TISSUE_NAME =
           case_when(Organ %in% "Muskel" ~ "Muscle",
                     Organ %in% "Lever" ~ "Liver"),
         STATION_NAME = Posisjon,                      # OR "= Posisjon", if given
         NIVA_CODE = 
           case_when(TISSUE_NAME %in% "Muscle" ~ "MU",
                     TISSUE_NAME %in% "Liver" ~ "LI")
  )


# Get last year's median
df_nifes_finalyear <- df_median %>% 
  filter(Parameter %in% sel_param & !is.na(Conc)) %>%
  group_by(STATION_NAME, Parameter, TISSUE_NAME, NIVA_CODE)  %>%
  mutate(Maxyear = max(Year)) %>%
  filter(Year == Maxyear) %>%
  # select(-LONGITUDE, -LATITUDE) %>%             # no LONGITUDE, LATITUDE this
  as.data.frame()

df_nifes_pos <- df_median %>% 
  filter(Parameter %in% sel_param & !is.na(Conc)) %>%
  group_by(STATION_NAME)  %>%
  summarize(
    LONGITUDE = median(LONGITUDE, na.rm = TRUE),
    LATITUDE = median(LATITUDE, na.rm = TRUE)
  )

df_nifes_finalyear <- 
  df_nifes_finalyear %>% left_join(df_nifes_pos)

xtabs(~Parameter + TISSUE_NAME, df_nifes_finalyear)  

```


## Make 'df_indicator_nifes'
```{r}

df_indicator_nifes <- tibble(
       PROJECT_ID = 9999,
       LATIN_NAME = "Gadus morhua",
       STATION_CODE = "9999", 
       STATION_NAME = df_nifes_finalyear$STATION_NAME,
       LONGITUDE = df_nifes_finalyear$LONGITUDE, 
       LATITUDE = df_nifes_finalyear$LATITUDE,
       PARAM = df_nifes_finalyear$Parameter,
       N = df_nifes_finalyear$N,
       Conc = df_nifes_finalyear$Conc,
       SPECIES_ID = 17,
       TISSUE_NAME = df_nifes_finalyear$TISSUE_NAME,
       NIVA_CODE = df_nifes_finalyear$NIVA_CODE
       )

# Add TISSUE_NAME to nifes_regression (for join below)
nifes_regression <- nifes_regression %>%
  mutate(TISSUE_NAME =
           case_when(TISSUE_NAME %in% "Muskel" ~ "Muscle",
                     TISSUE_NAME %in% "Lever" ~ "Liver",
                     TRUE ~ TISSUE_NAME),
         STATION_NAME = Posisjon
  )

# xtabs(~STATION_NAME, nifes_regression)
nrow(df_indicator_nifes)
df_indicator_nifes <- left_join(
  df_indicator_nifes, 
  nifes_regression[,c("PARAM", "TISSUE_NAME", "STATION_NAME", "trend")],
  by = c("STATION_NAME", "PARAM", "TISSUE_NAME"))
nrow(df_indicator_nifes)

# Check
# df_indicator_nifes %>% select(STATION_NAME, LATITUDE, Conc, NIVA_CODE, trend) %>% View()
# df_indicator %>% select(STATION_NAME, LATITUDE, Conc, NIVA_CODE, trend) %>% View()

```

## Combine NIVA and NIFES data
```{r}

df_indicator2 <- bind_rows(df_indicator, df_indicator_nifes)
df_indicator2 %>% head(2)
nrow(df_indicator2)  # 45

# Check liver
# xtabs(~STATION_NAME + PARAM, df_indicator2 %>% filter(TISSUE_NAME == "Lever"))  # 94

```

## Map data
```{r, fig.width=4, fig.height=5}

# Check that all stations have positions
sum(is.na(df_indicator2$LATITUDE))

df_indicator2 %>%
  count(STATION_NAME, LATITUDE,  LONGITUDE) %>%
  arrange(LATITUDE)

plot(LATITUDE ~ LONGITUDE, df_indicator2)
maps::map(regions = "Norway", add = TRUE, col = "brown")


```

## Proref, EQS and mattrygghet
```{r}

# Get PROREF (Q95) and EQS
df_q95 <- data_xl_sel[data_xl_sel$Basis %in% "WW", 
                       c("Parameter Code", "Species", "Tissue", "PROREF", "EQS-threshold this year")] %>%
  rename(PARAM = `Parameter Code`, 
         LATIN_NAME = Species, 
         TISSUE_NAME = Tissue) %>%
  group_by(PARAM, LATIN_NAME, TISSUE_NAME) %>%
  summarise(PROREF = first(PROREF), EQS_threshold = first(`EQS-threshold this year`))

# Check
# df_q95 %>% filter(PARAM == "HG")

# add PROREF and EQS thresholds
nrow(df_indicator2) # 94
df_indicator2 <- left_join(df_indicator2, 
                           df_q95,
                           by = c("LATIN_NAME", "TISSUE_NAME", "PARAM"))
nrow(df_indicator2) # 94

# change EQS threshold for Hg (by mail from Norman; from 1.00 to 0.67)
sel <- df_indicator2$PARAM %in% "CB_S7"; sum(sel) 
df_indicator2$EQS_threshold[sel] <- 0.67

# Classes changed - original classes were: c(-999999,1,2,5,10,20,999999)
df_indicator2$KLASSE <- with(df_indicator2, 
                             cut(Conc/PROREF, breaks = c(-999999,1,2,10,999999), 
                                 right = FALSE, labels = FALSE)
                             )

xtabs(~KLASSE + PARAM, df_indicator2)
#       PARAM
# KLASSE BDE6S CB_S7 CD DDEPP HCB HG PB
#      1     7     7  8     1   1  3  8
#      2     0     1  0     0   0  1  0
#      3     0     0  0     0   0  4  0
     

# EQS classe
df_indicator2$EQS <- with(df_indicator2, 
                          cut(Conc/EQS_threshold, breaks = c(-999999,1,999999), 
                              right = FALSE, labels = FALSE)
                          )

xtabs(~addNA(EQS) + PARAM, df_indicator2)
#           PARAM
# addNA(EQS) BDE6S CB_S7 CD DDEPP HCB HG PB
#       1        0     0  0     1   1  0  0
#       2        7     8  0     0   0  8  0
#       <NA>     0     0 11     0   0  3 11
      



```

## Add SPECIES_ID
```{r}

df_indicator2 <- df_indicator2 %>%
  mutate(SPECIES_ID = case_when(
    LATIN_NAME %in% "Mytilus edulis" ~ 26,
    LATIN_NAME %in% "Gadus morhua" ~ 17)
  )

```

## Add food limit  
1 = below limit, 2 = above limit
```{r}

nrow(df_indicator2)
df_indicator2 <- left_join(
  df_indicator2, 
  subset(df_limits, select = c(PARAM, LATIN_NAME, NIVA_CODE, Mattrygghet)), 
  by = c("PARAM", "LATIN_NAME", "NIVA_CODE"))
nrow(df_indicator2)
# 96

df_indicator2$Mattrygghet <- cut(with(df_indicator2, Conc/Mattrygghet),
                                 breaks = c(-999999,1,999999), 
                                 right = FALSE, labels = FALSE)

xtabs(~addNA(Mattrygghet) + PARAM + TISSUE_NAME, df_indicator2)

# tail(df_indicator2)

```

## Save and export  
* On 'data': save with all variables   
* On 'Data_export': pick selected variables  
```{r}
# Order logically
df_indicator2 <- df_indicator2 %>%
  filter(LATIN_NAME %in% "Gadus morhua") %>% 
  arrange(PARAM,LONGITUDE,LATITUDE)

# Save with all variables
saveRDS(df_indicator2, file = "Data/14_df_indicator2_cod_ver01 (2021.rds")

# Save with selected variables
vars <- c("PROJECT_ID", "STATION_CODE", "STATION_NAME", 
  "LONGITUDE", "LATITUDE", "PARAM", "N", "KLASSE", 
  "SPECIES_ID", "trend", "NIVA_CODE", "EQS", "Mattrygghet")

# DOnt include HG in liver
sel <- with(df_indicator2, !(PARAM == "HG" & TISSUE_NAME == "Lever"))
sum(sel)
mean(sel)


# setwd("H:/Documents/seksjon 212/Indikator 2018/Analyse")


write_safe <- function(data, filename, save_function, overwrite = FALSE, ...){
  # Csv
  if (overwrite | !file.exists(filename)){
    save_function(data, file = filename, ...)
    cat("Data written to file", filename, "\n")
  } else {
    cat("File", filename, "exists, data not saved.\nSet 'overwrite = TRUE' to overwrite file anyway.\n")
  }
}

# Csv
write_safe(df_indicator2[sel,vars], filename = fn_csv, save_function = write.csv,
           overwrite = overwrite, row.names = FALSE, na = "")

# Excel
write_safe(df_indicator2[sel,vars], filename = fn_xlsx, save_function = openxlsx::write.xlsx,
           overwrite = overwrite)




# df_indicator2 <- read.csv2(file = "Data_export/GaduMor_2018_withNIFES_ver1.csv")

```

## Checks
### Plots, metals
trend: 0 = no trend calculated, 1 = zero time trend, 2 = up, 3 = down  
```{r}

df <- df_indicator2 %>%
  filter(!(PARAM == "HG" & TISSUE_NAME == "Lever")) %>%
  mutate(Time_trend =
           case_when(trend %in% 0 ~ "Not calc.",
                     trend %in% 1 ~ "No trend",
                     trend %in% 2 ~ "Down",
                     trend %in% 3 ~ "Up")
         ) %>%
  mutate(Time_trend = factor(Time_trend, levels = c("Not calc.", "No trend", "Down", "Up")),
         Mattrygghet = factor(Mattrygghet, levels = 1:2),
         EQS = factor(EQS, levels = 1:2))

# Fill = trend
gg1 <- df %>%
  filter(PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Time_trend)) + 
  geom_col() + 
  scale_fill_manual(values = c("grey30", "orange", "green", "red"), drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

# Fill = Mattrygghet
gg2 <- df %>%
  filter(PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Mattrygghet)) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red"), na.value = "grey30", drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

# Fill = EQS
gg3 <- df %>%
  filter(PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = EQS)) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red"), na.value = "grey30", drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

# ggsave()

gg1
gg2
gg3
```

### Plots, non-metals
```{r}

gg4 <- df %>%
  filter(!PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Time_trend)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("grey30", "orange", "green", "red"), drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))


# Fill = Mattrygghet
gg5 <- df %>%
  filter(!PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Mattrygghet)) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red"), na.value = "grey30", drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

# Fill = EQS
gg6 <- df %>%
  filter(!PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = EQS)) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red"), na.value = "grey30", drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

gg4
gg5
gg6

```

## Compare with last time  
```{r}

df_compare <- df_indicator2 %>%
  select(PARAM, NIVA_CODE, STATION_CODE, KLASSE) %>%
  left_join(
    df_2017 %>% 
      select(PARAM, NIVA_CODE, STATION_CODE, KLASSE) %>%
      rename(KLASSE_2017 = KLASSE)
  )

xtabs(~KLASSE_2017 + KLASSE + PARAM, df_compare)

```


## Map data  
```{r, fig.width=6, fig.height=4}

param <- "CB_S7"
df %>% 
  filter(PARAM %in% param) %>%
  ggplot() +
    annotation_map(simple_map, fill = "olivedrab3") +
    geom_point(aes(LONGITUDE, LATITUDE, color = Mattrygghet)) +
    scale_color_manual(values = c("green3","red2")) +
  coord_equal(ratio = 2) +
  labs(title = paste0(param, ", Mattrygghet"))
  

df_2017 %>% 
  filter(PARAM %in% "CB_S7") %>%
  ggplot() +
    annotation_map(simple_map, fill = "olivedrab3") +
    geom_point(aes(LONGITUDE, LATITUDE, color = factor(EQS))) +
    scale_color_manual(values = c("green3","red2")) +
  coord_equal(ratio = 2)


```

