= ---
title: "Combine NIVA and NIFES data 2022 - Norwegian Sea"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
---

Combining medians from   
* NIFES data (script 11) - special script for 2022_NorwSea
* Trends from NIFES data (script 12) - special script for 2022_NorwSea   
* Medians and trends from NIVA (script 13) - same script as 2022 North Sea (set 'area' to "Norweegian Sea')  
* In addition to merging the results, limit values are also added in this script    
  
The resulting summary data (one line per time series, i.e. station/substance) are saved as csv and excel files. These are also copied to `K:\Avdeling\Mar\NOG\JMGdata\Kart\Indikatorer\2020\Grunnlagsdata Norskehavet`      



## Packages + functions  
```{r}

library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)
library(stringr)


# File name for export
fn <- "Data_export/GaduMor_2022data_ver01"
fn_csv <- paste0(fn, ".csv")
fn_xlsx <- paste0(fn, ".xlsx")
overwrite <- TRUE
#overwrite <- FALSE

# Background map
library(mapdata)
simple_map <- map_data("worldHires", "Norway") %>% fortify()

```

## Data
Indicators from NIVA serves as the starting point for combining the data  
```{r}

# NIVA & NIFES and indicators
df_indicator <- readRDS("Data/13_df_indicator_NIVA_only (2022).rds")

# NIFES medians
df_median <- readRDS("Data/13_df_indicator_median (2022).rds") %>%
  as.data.frame()

df_median_2022 <- df_median %>% 
  filter(MYEAR == 2022)

# Food and EQS limits
df_limits <- read_excel("Input_data/Grenseverdier_fra_Sylvia.xlsx")
  

# Parameters
sel_param <- c("CD", "HG", "PB", "HCB", "DDEPP", "CB_S7", "BDE6S")

niva_stations <- readRDS("Input_data/NIVA_data_stations.rds")
proref <- readRDS("Data/lookup_proref.rds")

proref1 <- proref %>% 
  rename(PARAM = `Parameter Code`)

```

### a. Add "station code"  

```{r}

df_nifes_cod <- readRDS("Data/11_df_nifes_cod (2022).rds")
#fix concentration
df_nifes_cod1 <- df_nifes_cod %>% 
  mutate( Conc = sub("<", "", Conc_chr, fixed = TRUE) %>% 
            sub("*", "", ., fixed = TRUE) %>%
            as.numeric(),
          Flag = ifelse(grepl("<", Conc_chr), "<", ""),
          Comment = ifelse(grepl("*", Conc_chr, fixed = TRUE), "*", "")
    )

df_nifes_cod1 <- df_nifes_cod1 %>% 
  filter(Parameter %in% c("Cd", "Pb", "Hg", "PCB-118", "Sum PCB 7", 
                          "Hexachlorobenzene (HCB)", "pp-DDE", "PBDE 47", "BDE6S")) %>% 
  filter(Year== 2022) %>% 
  group_by(Posisjon, Parameter, Organ) %>%
  mutate(Conc_median = median(Conc))

df_nifes_cod1 <- df_nifes_cod1 %>% 
  select(-Conc, -Parameter, -Organ)

df_nifes_cod1 <- df_nifes_cod1 %>% 
 rename("Conc" = Conc_median) %>% 
  select(Conc, PARAM, STATION_CODE)

# fix station names
df_nifes_stations <- df_nifes_cod %>%
  distinct(LONGITUDE, LATITUDE) %>%
  mutate(STATION_CODE = paste0("HI_Barents", 1:n()))
if (!"STATION_CODE" %in% names(df_nifes_cod)){
  df_nifes_cod <- df_nifes_cod %>%
    left_join(df_nifes_stations, by = c("LATITUDE", "LONGITUDE"))
}
 # Find less-thans and make Flag variable

```


## Add food limit  
1 = below limit, 2 = above limit
```{r}
df_indicator2 <- df_indicator %>%
  full_join(df_median_2022, multiple ="any") %>% 
  left_join(niva_stations, multiple ="any") %>% 
  left_join(niva_stations, multiple ="any") %>% 
  left_join(df_nifes_stations, multiple ="any") %>% 
  left_join(df_nifes_cod1, multiple ="any") %>% 
  left_join(proref1, multiple ="any")
  
  
df_indicator2 <- df_indicator2 %>% 
  select(-MYEAR, -VALUE_lo, -VALUE_up, -Posisjon, -Parameter, -Organ)
### adding Latin_Name according to 
#LATIN_NAME is "Gadus morhua" for cod, i.e. all NIVA stations with "B" in the station code and all NIFES/HI stations, and "Mytilus edulis" (blue mussel) for the rest
# NIVA_CODE is "LI" (liver) for NIVA cod, except for HG for which it is "MU"
#"SB" (soft body) for the blue mussel stations. For the NIFES data (cod) it's the same as NIVA cod I think (should say in the original excel)
# add latine name
df_indicator2 <- df_indicator2 %>% mutate(
  LATIN_NAME= 
    case_when(
      grepl("B",STATION_CODE)~ "Gadus morhua",
      grepl("HI_Barents", STATION_CODE) ~ "Gadus morhua",
      TRUE ~ "Mytilus edulis"
      ))
#add nIVA code
df_indicator2 <- df_indicator2 %>% mutate(
  NIVA_CODE = 
    case_when(
      LATIN_NAME %in% "Mytilus edulis" ~ "SB",
      LATIN_NAME %in% "Gadus morhua" & PARAM == "HG" ~ "MU",
      TRUE ~ "LI"
    ))
# add tissue name
df_indicator2 <- df_indicator2 %>% mutate(
  TISSUE_NAME = 
    case_when(
      LATIN_NAME %in% "Mytilus edulis" ~ "Soft body",
      LATIN_NAME %in% "Gadus morhua" & PARAM == "HG" ~ "Muscle",
      TRUE ~ "Liver"
    ))

#add class
# Note: the values I gave are for cod only - blue mussels should not be included at all (sorry for that! I had forgotten this)
# KLASSE is found by comparing median concentration (last year) with PROREF
df_indicator2 <- df_indicator2 %>% mutate(
  CLASS = case_when(
    Conc/PROREF < 1 ~ 1,
    Conc/PROREF < 2 ~ 2,
    Conc/PROREF < 10 ~ 3,
    Conc/PROREF >= 10 ~ 4
  ))

nrow(df_indicator2)
df_indicator2 <- left_join(
  df_indicator2, 
  subset(df_limits, select = c(PARAM, LATIN_NAME, NIVA_CODE, Mattrygghet)), 
  by = c("PARAM", "LATIN_NAME", "NIVA_CODE"))
nrow(df_indicator2)

df_indicator2$Mattrygghet <- cut(with(df_indicator2, Conc/Mattrygghet),
                                 breaks = c(-999999,1,999999), 
                                 right = FALSE, labels = FALSE)

xtabs(~addNA(Mattrygghet) + PARAM + TISSUE_NAME, df_indicator2)

#tail(df_indicator2)

```

## Save and export  
* On 'data': save with all variables   
* On 'Data_export': pick selected variables  
```{r}
# Order logically
df_indicator2 <- df_indicator2 %>%
  filter(LATIN_NAME %in% "Gadus morhua") %>% 
  arrange(PARAM,LONGITUDE,LATITUDE)

# Save with all variables
saveRDS(df_indicator2, file = "Data/14_df_indicator2_cod_ver01 (2022.rds")

# Save with selected variables (deleted PROJECT_ID, N and SPECIES_ID)
vars <- c("STATION_CODE", "STATION_NAME", "LONGITUDE", "LATITUDE", "PARAM", "CLASS", "trend", "NIVA_CODE", "EQS", "Mattrygghet")

# DOnt include HG in liver
sel <- with(df_indicator2, !(PARAM == "HG" & TISSUE_NAME == "Liver"))
sum(sel)
mean(sel)

write_safe <- function(data, filename, save_function, overwrite = FALSE, ...){
  # Csv
  if (overwrite | !file.exists(filename)){
    save_function(data, file = filename, ...)
    cat("Data written to file", filename, "\n")
  } else {
    cat("File", filename, "exists, data not saved.\nSet 'overwrite = TRUE' to overwrite file anyway.\n")
  }
}

# Csv
write_safe(df_indicator2[sel,vars], filename = fn_csv, save_function = write.csv,
           overwrite = overwrite, row.names = FALSE, na = "")

# Excel
write_safe(df_indicator2[sel,vars], filename = fn_xlsx, save_function = openxlsx::write.xlsx,
           overwrite = overwrite)




# df_indicator2 <- read.csv2(file = "Data_export/GaduMor_2018_withNIFES_ver1.csv")

```

## Checks
### Plots, metals
trend: 0 = no trend calculated, 1 = zero time trend, 2 = up, 3 = down  
```{r}

df <- df_indicator2 %>%
  filter(!(PARAM == "HG" & TISSUE_NAME == "Liver")) %>%
  mutate(Time_trend =
           case_when(trend %in% 0 ~ "Not calc.",
                     trend %in% 1 ~ "No trend",
                     trend %in% 2 ~ "Down",
                     trend %in% 3 ~ "Up")) %>%
  mutate(Time_trend = factor(Time_trend, levels = c("Not calc.", "No trend", "Down", "Up")),
         Mattrygghet = factor(Mattrygghet, levels = 1:2),
         EQS = factor(EQS, levels = 1:2))

# Fill = trend
gg1 <- df %>%
  filter(PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Time_trend)) + 
  geom_col() + 
  scale_fill_manual(values = c("grey30", "orange", "green", "red"), drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

# Fill = Mattrygghet
gg2 <- df %>%
  filter(PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Mattrygghet)) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red"), na.value = "grey30", drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

# Fill = EQS
gg3 <- df %>%
  filter(PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = EQS)) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red"), na.value = "grey30", drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

# ggsave()

gg1
gg2
gg3
```

### Plots, non-metals
```{r}

gg4 <- df %>%
  filter(!PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Time_trend)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("grey30", "orange", "green", "red"), drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))


# Fill = Mattrygghet
gg5 <- df %>%
  filter(!PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = Mattrygghet)) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red"), na.value = "grey30", drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

# Fill = EQS
gg6 <- df %>%
  filter(!PARAM %in% c("HG", "PB", "CD")) %>%
  ggplot(aes(STATION_NAME, Conc, fill = EQS)) + 
  geom_col() + 
  scale_fill_manual(values = c("green", "red"), na.value = "grey30", drop = FALSE) +
  facet_wrap(~paste(PARAM,TISSUE_NAME), scales = "free", nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))

gg4
gg5
gg6

```

## Compare with last time  
```{r}

df_compare <- df_indicator2 %>%
  select(PARAM, NIVA_CODE, STATION_CODE, KLASSE) %>%
  left_join(
    df_2017 %>% 
      select(PARAM, NIVA_CODE, STATION_CODE, KLASSE) %>%
      rename(KLASSE_2017 = KLASSE)
  )

xtabs(~KLASSE_2017 + KLASSE + PARAM, df_compare)

```


## Map data  
```{r, fig.width=6, fig.height=4}

param <- "CB_S7"
df %>% 
  filter(PARAM %in% param) %>%
  ggplot() +
    annotation_map(simple_map, fill = "olivedrab3") +
    geom_point(aes(LONGITUDE, LATITUDE, color = Mattrygghet)) +
    scale_color_manual(values = c("green3","red2")) +
  coord_equal(ratio = 2) +
  labs(title = paste0(param, ", Mattrygghet"))
  

df_2017 %>% 
  filter(PARAM %in% "CB_S7") %>%
  ggplot() +
    annotation_map(simple_map, fill = "olivedrab3") +
    geom_point(aes(LONGITUDE, LATITUDE, color = factor(EQS))) +
    scale_color_manual(values = c("green3","red2")) +
  coord_equal(ratio = 2)


```

