---
title: "15b_Final_results_fix"
format: html
---

Checking and fixing errors iin PBDE6 for HI's Barents Sea stations  

## Packages + functions  
```{r}

library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)
library(stringr)
library(knitr)    # for kable

# for JAGS (interval regression)  
library(R2jags)
library(rjags)
library(purrr)
source("interval_lm_qi.R")  # function 'int_linear_qi'  

pars_bde6s <- c("PBDE 28","PBDE 47","PBDE 99","PBDE 100","PBDE 153","PBDE 154")

```
## Checking  

### Check summary data sent by Sylvia   

- attachement in mail from Merete S 20-3-2024   

```{r}

fn <- "Input_data/Data til figurer Barentshav 2023.xlsx"  
excel_sheets(fn)

df_sylvia_headers <- read_xlsx(fn, sheet = "Torsk, lever")
df_sylvia_headers2 <- read_xlsx(fn, sheet = "Torsk, lever", col_names = FALSE)
df_sylvia <- read_xlsx(fn, sheet = "Torsk, lever", skip = 3, col_names = FALSE)

names(df_sylvia) <- names(df_sylvia_headers)
df_sylvia_headers2[1:3, 13:16]
names(df_sylvia)[13:16] <- paste0("PBDE_", c("N", "mean", "min", "max"))
  
ggplot(df_sylvia, aes(År, PBDE_mean)) + 
  geom_point() + 
  geom_line() + 
  theme_minimal()

```


### Check delivered data   

```{r}
  
df_deliver <- read_xlsx("Data_export/2022 report Barents Sea/Indicators_Barents_Sea_2022_ver6_for Sylvia.xlsx")
table(df_deliver$PARAM)
ggplot(df_deliver %>% filter(PARAM == "BDE6S"), aes(STATION_CODE, Conc)) + 
  geom_col()

df_deliver <- read_xlsx("Data_export/2022 report Barents Sea/Indicators_Barents_Sea_2022_ver6.xlsx")
df_deliver <- read_xlsx("Data_export/2022 report Barents Sea/Indicators_Barents_Sea_2022_ver3.xlsx")
# table(df_deliver$PARAM)
ggplot(df_deliver %>% filter(PARAM == "BDE6S"), aes(STATION_CODE, Conc)) + 
  geom_col()

```

###  Data used in script 15    

- data created by script 11  

```{r}

#
# Raw data for HI (NIFES) cod    
# - includes CB_S7 and BDE6S
#
dat_hi <- readRDS("Data/11_df_nifes_cod (2022).rds")
xtabs(~PARAM + STATION_CODE, dat_hi)

# names(dat_hi)

```

#### Check sum PBDE  

```{r}

ggplot(dat_hi %>% filter(PARAM == "BDE6S"), aes(Year, Conc, color = Flag)) + 
  geom_jitter(width = 0.2) + 
  facet_wrap(vars(STATION_CODE))

```

### Original data supplied by Sylvia   

- from the start of script 11  

```{r}

fn <- "Input_data/POPs torskelever Bhav 2006-2022 inkl. PBDE HI.xlsx"
# excel_sheets(fn)
df_nifes_cod1 <- read_excel(fn, col_types = "text") %>%
  rename(Prøvenr. = `Prøve nr.`)  %>%     # fix this column name (equal to dat1)
  mutate(Year = as.numeric(År))

# df_nifes_cod1[,c("Year", pars_bde6s)]

```

#### Add "station code"  

- Based on the dataset created at the end of script 11

```{r}

table(addNA(df_nifes_cod1$Posisjon))
table(addNA(df_nifes_cod1$Område))

# dat_hi is the data made by script 11
df_nifes_stations <- dat_hi %>%
  distinct(Posisjon, STATION_CODE)

df_nifes_cod2 <- df_nifes_cod1 %>%
  left_join(df_nifes_stations, by = c("Posisjon"))

dat_stations

```



#### Check for units different than ng/g  

- NOTE: visual check

```{r}

columns_unit <- which(names(dat3) %in% pars_bde6s) + 1
dat3[,columns_unit]

# Check for units different than ng/g  
for (i in columns_unit)
  print(table(dat3[[i]]))

```
### BDE data  

#### Select columns, reformat to get data   
```{r}

df_nifes_bde_wide <- df_nifes_cod2[,c("Year", "Posisjon", "STATION_CODE", "Prøvenr.", pars_bde6s)]

df_nifes_bde <- df_nifes_bde_wide %>%
  pivot_longer(-c(Year, Posisjon, STATION_CODE, Prøvenr.), names_to = "Parameter", values_to = "Conc_chr") %>%
  filter(!is.na(Conc_chr)) %>%
  mutate(
    Conc = sub("<", "", Conc_chr, fixed = TRUE) %>% 
      sub("*", "", ., fixed = TRUE) %>%
      as.numeric(),
    Flag = ifelse(grepl("<", Conc_chr), "<", ""),
    Comment = ifelse(grepl("*", Conc_chr, fixed = TRUE), "*", "")
  )

```

#### Test plot by parameter  

```{r}

ggplot(df_nifes_bde, aes(Year, Conc, color = Flag)) + 
  geom_jitter(width = 0.2) + 
  facet_grid(vars(Parameter), vars(STATION_CODE))

```
#### Overl LOQ  
```{r}

df_nifes_bde %>%
  group_by(Year, Parameter, STATION_CODE) %>%
  summarise(
    Above_LOQ = 100*mean(is.na(Flag) | Flag %in% "")
  ) %>%
  ggplot(aes(Year, Above_LOQ)) + 
  geom_point() +
  geom_line() +
  facet_grid(vars(Parameter), vars(STATION_CODE))

```

#### Sum BDE6S  

```{r}

# sel <- is.na(df_nifes_bde$Flag); sum(sel)
# View(df_nifes_bde[sel,])

table(addNA(df_nifes_bde$Flag))

df_nifes_bde6S <- df_nifes_bde %>%
  group_by(Year, Posisjon, STATION_CODE, Prøvenr.) %>%
  summarise(
    Above_LOQ = sum(is.na(Flag) | Flag %in% ""),
    Conc_incl_LOQ = sum(Conc),
    Conc_excl_LOQ = sum(Conc[is.na(Flag) | Flag %in% ""])
  )

table(df_nifes_bde6S$Conc_incl_LOQ - df_nifes_bde6S$Conc_excl_LOQ > 0)
```
#### test plot sum  

```{r}

ggplot(df_nifes_bde6S, aes(Year, Conc, color = Flag)) + 
  geom_jitter(width = 0.2) + 
  facet_grid(vars(Parameter), vars(STATION_CODE))

```


```{r}
