---
title: "21_Combine_maps"
author: "DHJ"
date: "30 9 2021"
output: html_document
---


**Adding inset of Tromsø to maps of Norwegian Sea south**   

- Read images of two maps  
- Crops one of them, adds text and border
- Adds this as an inset to the other  
- Also see test file in 'R_test' project



[Manual for the magick package](https://cran.r-project.org/web/packages/magick/vignettes/intro.html)

```{r}

library(magick) 
library(dplyr)
library(purrr)

```

## Read files   
- Norwegian Sea, south and north parts  
- Size = 945 x 945   
```{r}

# Original
folder <- "K:/Avdeling/Mar/NOG/JMGdata/Kart/Indikatorer/2018/Figurer rev sept 2021"

dir(folder)

# Norwegian Sea, south 
fn <- paste0(folder, "/", "GADUMOR_BDE6S_M.png")
map_south <- image_read(fn)

# Norwegian Sea, north 
fn <- paste0(folder, "/", "GADUMOR_BDE6S_N.png")
map_north <- image_read(fn)

```

## Test procedure  

### Crop image  
```{r}

# example of "geometry" (check ?geometry)
# "500x300+10+20" – Crop image to 500 by 300 at position 10,20 (position measured from TOP LEFT)
# Add '!' to keep aspect

# Use the following command
#   image_crop(map_north, "400x400+50+50!")
# in the console, repeatedly (with different geometries), to find a good crop  

inset_1 <- image_crop(map_north, "200x200+260+420!")  
inset_1

```

### Add text and border   
```{r}

inset_2 <- image_annotate(inset_1, "Tromsø", size = 35, gravity = "north")

### Add border  
inset_3 <- image_border(inset_2, "black", "2x2")

inset_3

```

### Put inset into the southern map
```{r}

map_comb <- image_composite(map_south, inset_3, offset = "+650+70")
map_comb

```



## Use function  

### Function   
* Reads files ending in "_N" and "_M", making insets  
* New figures are saved with "_comb" in name  
```{r}

make_combined_map <- function(txt, folder){
  
  # Norwegian Sea, south 
  fn <- paste0(folder, "/", txt, "_M.png")
  map_south <- image_read(fn)
  
  # Norwegian Sea, north 
  fn <- paste0(folder, "/", txt, "_N.png")
  map_north <- image_read(fn)
  
  # Crop northern map
  inset_1 <- image_crop(map_north, "200x200+260+420!")  
  
  # Add text  
  inset_2 <- image_annotate(inset_1, "Tromsø", size = 35, gravity = "north")
  
  # Add border  
  inset_3 <- image_border(inset_2, "black", "2x2")
  
  # Put inset into the southern map
  # Use image_composite  
  map_comb <- image_composite(map_south, inset_3, offset = "+650+70")
  
  # Save
  fn <- paste0(folder, "/", txt, "_comb.png")
  image_write(map_comb, path = fn, format = "png")
  
}

make_combined_map("GADUMOR_BDE6S", "11_magick")

```

### Use in original folder  
```{r}

# Original
folder <- "K:/Avdeling/Mar/NOG/JMGdata/Kart/Indikatorer/2018/Figurer rev sept 2021"

# Used to make 'map_txt' below:
# grep("_M", dir(folder), value = TRUE) %>% 
#   sub("_M.png", "", ., fixed = TRUE) %>%
#   dput()

map_txt <- c("GADUMOR_BDE6S", "GADUMOR_CBS7", "GADUMOR_CD", "GADUMOR_DDEPP", 
             "GADUMOR_HCB", "GADUMOR_HG", "GADUMOR_PB")

# Uncomment next line to run - this actually makes/overwrites the '_comb' picture files  
# walk(map_txt, make_combined_map, folder = folder)


```

## Make an inset out of 'Tromsø' in the northern map   

### Crop image  
```{r}

# example of "geometry" (check ?geometry)
# "500x300+10+20" – Crop image to 500 by 300 at position 10,20 (position measured from TOP LEFT)
# Add '!' to keep aspect

# Use just sentence such as this
# image_crop(map_north, "400x400+50+50!")
# In the console, repeatedly, to find a good crop  

inset_1 <- image_crop(map_north, "200x200+260+420!")  

### Add text  
inset_2 <- image_annotate(inset_1, "Tromsø", size = 35, gravity = "north")

### Add border  
inset_3 <- image_border(inset_2, "black", "2x2")

## Put inset into the southern map
### Use image_composite  
map_comb <- image_composite(map_south, inset_3, offset = "+650+70")

fn <- paste0(folder, "/", "GADUMOR_BDE6S_comb.png")

image_write(map_comb, path = fn, format = "png")

```